---
title: "Pairwise DE figure work"
author: "Cat Brown"
output:
  html_notebook:
    toc: yes
    toc_float: yes
---

# Load libraries
```{r, message=FALSE}
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(tidyverse)
library(writexl)
library(readxl)
library(cowplot)
library(svglite)
library(pheatmap)
library(ggVennDiagram)
``` 

# Read in Atlas
```{r}
atlas <- readRDS("C:/Users/cat14/OneDrive - Johns Hopkins/Farber_Lab/Basecaller contents/atlas_paper/251119_atlas_v2_with_cell_cycle_score_info.rds")
```

# Load pairwise DE results
```{r}
pairwise_full <- read.csv("C:/Users/cat14/OneDrive - Johns Hopkins/Farber_Lab/Basecaller contents/atlas_paper/260129_short_cell_type_cluster_v2_0_pairwise_DE_from_250113.csv")
```

# Filter to just comparisons to unfed
```{r}
pairwise_full <- filter(pairwise_full, time %in% c("t1-15HF.t0-UF", "t2-30HF.t0-UF", "t3-60HF.t0-UF", "t4-120HF.t0-UF", "t5-240HF.t0-UF"))
```


# Get unique pairwise genes
```{r}
genes <- pairwise_full %>% select("gene") %>% unique() %>% unlist()
```

# Get average expression levels for pairwise genes in all clusters
```{r}
avg_exp <- AverageExpression(atlas, group.by = c("short_cell_type_cluster", "sample_name"), assay = "RNA", features = genes)
```
# Restructure avg_exp so that I can combine with LRT
```{r}
avg_exp_df <- as.data.frame(avg_exp[["RNA"]]) %>% 
  rownames_to_column(var = "gene") %>%
  pivot_longer(cols = 2:4023, values_to = "expression", names_to = "cluster_sample") %>%
  separate_wider_delim(cols = "cluster_sample", delim = "_", names = c("cluster", "sample")) %>%
  pivot_wider(names_from = "sample", values_from = "expression")
```


# Combine to filter the average expression to relvant rows
```{r}
pairwise_tags <- select(pairwise_full, c("gene", "short_cell_type_cluster"))
heatmap_data <- left_join(pairwise_tags, avg_exp_df, by = c("gene" = "gene", "short_cell_type_cluster" = "cluster"))
```

# Create matrix with gene_cluster as rownames and expression as the data
```{r}
heatmap_matrix <- tidyr::unite(heatmap_data, "gene":"short_cell_type_cluster", col="gene_celltype", sep = "_") %>% unique() %>% relocate(`t3-sample9-60HF`, .after = `t2-sample8-30HF`)  %>% remove_rownames() %>% column_to_rownames(var = "gene_celltype") %>% as.matrix()
```

```{r}
time_colors <- c("unfed" = "hotpink1",  "15min" = "orange", "30min" = "yellow2", "1h" = "lightgreen", "2h" = "steelblue1", "4h" = "mediumpurple1")
sample_times_info <- c("unfed", "unfed", "unfed", "15min", "15min", "15min", "30min", "30min", "1h", "1h", "1h", "2h", "2h", "4h", "4h")
sample_times <- data.frame(condition = sample_times_info, row.names = colnames(heatmap_matrix))
```

# Create tagging system for DE times
```{r}
DE_time_tags <- pairwise_full %>% dplyr::select(c("gene","short_cell_type_cluster", "time")) %>%
  tidyr::unite("gene":"short_cell_type_cluster", col="gene_celltype", sep = "_") %>%
  unique() %>% group_by(gene_celltype)
DE_time_tags$time <- DE_time_tags$time %>% str_remove(".t0-UF")
DE_time_tags <- DE_time_tags %>% nest("time") %>% rowwise() %>%
  mutate(times = list(unlist(data, use.names = FALSE))) %>% 
  mutate(times = paste(times, collapse = ", ")) %>%
  select(!"data") %>% column_to_rownames(var = "gene_celltype")
```

```{r}
comparison_colors <- c("t1-15HF" = "khaki2", 
                       "t1-15HF, t2-30HF" = "maroon", 
                       "t1-15HF, t2-30HF, t3-60HF" = "orchid1", 
                       "t1-15HF, t2-30HF, t3-60HF, t4-120HF" = "deeppink1", 
                       "t1-15HF, t2-30HF, t3-60HF, t4-120HF, t5-240HF" = "blue1", 
                       "t1-15HF, t2-30HF, t4-120HF" = "steelblue4",
 "t1-15HF, t3-60HF" = "darkturquoise",
 "t1-15HF, t3-60HF, t4-120HF" = "green1",
 "t1-15HF, t3-60HF, t4-120HF, t5-240HF" = "yellow4",
 "t1-15HF, t3-60HF, t5-240HF" = "yellow3", 
 "t1-15HF, t4-120HF" = "darkorange4",
 "t1-15HF, t4-120HF, t5-240HF" = "brown", 
 "t1-15HF, t5-240HF" = "#6A3D9A", 
 "t2-30HF" = "dodgerblue2", 
 "t2-30HF, t3-60HF" = "#E31A1C", 
 "t2-30HF, t3-60HF, t4-120HF" = "green4", 
 "t2-30HF, t4-120HF" = "#FF7F00", 
 "t2-30HF, t5-240HF" = "black", 
 "t3-60HF" = "gold1", 
 "t3-60HF, t4-120HF" = "skyblue2", 
 "t3-60HF, t4-120HF, t5-240HF" = "#FB9A99", 
 "t3-60HF, t5-240HF" = "palegreen2", 
 "t4-120HF" = "#CAB2D6", 
 "t4-120HF, t5-240HF" = "#FDBF6F", 
 "t5-240HF" = "gray70")
```

# Make heatmap
```{r, fig.height=10, fig.width=4.75}
h1 <- pheatmap(heatmap_matrix, scale = "row",  cluster_cols = FALSE, show_rownames = FALSE, annotation_col = sample_times, annotation_colors = list(condition = time_colors, times = comparison_colors), annotation_row = DE_time_tags, fontsize = 6)
```
```{r}
save_plot("260624_pairwise_DE_expression_heatmap.png", h1, base_height = 10, base_width = 4.75)
save_plot("260624_pairwise_DE_expression_heatmap.svg", h1, base_height = 10, base_width = 4.75)
```

# VennDiagram of genes and times
## Reshape data
```{r}
DE_Venn_df <- pairwise_full %>% select(c("gene", "time")) %>% unique() %>%
  group_by(time) %>% nest(gene) %>% column_to_rownames(var = "time")
DE_Venn_list <- NULL
for (t in rownames(DE_Venn_df)) {
  DE_Venn_list <- c(DE_Venn_list, list(unlist(DE_Venn_df[t, "data"], use.names = FALSE))) # unlist turns it into just a character vector so list is required for it to append correctly. Omitting both results in a list with too many levels
}
names(DE_Venn_list) <- rownames(DE_Venn_df)
DE_Venn_list <- DE_Venn_list[c("t1-15HF.t0-UF", "t2-30HF.t0-UF", "t3-60HF.t0-UF", "t4-120HF.t0-UF", "t5-240HF.t0-UF")]
```
```{r, fig.width=3, fig.height=2}
DE_Venn_plot <- ggVennDiagram(DE_Venn_list, label_alpha = 0, 
                                  label = "count",
                                  label_size = 3, edge_size = 0.5) + 
  scale_fill_gradient(low = "#FFFFFF", high = "#FF66CC", trans = "sqrt") +
  theme(legend.position = "none", text = element_text(family = "sans", size = 6)) +
  scale_x_continuous(expand = expansion(mult = .4))
DE_Venn_plot
```
I bet I can move these in Illustrator
```{r}
save_plot("260624_pairwise_gene_overlap_venn.svg", DE_Venn_plot, base_width = 3, base_height = 2)
```
# Get genes per cluster
```{r}
times <- c("t1-15HF.t0-UF", "t2-30HF.t0-UF", "t3-60HF.t0-UF", "t4-120HF.t0-UF", "t5-240HF.t0-UF", "overall")
celltypes <- unique(atlas$short_cell_type_cluster)
celltypes <- str_replace_all(celltypes, "_", "-")
genes_per_cluster <- data.frame(matrix(ncol=length(times), nrow = length(celltypes)))
colnames(genes_per_cluster) <- times
row.names(genes_per_cluster) <- celltypes

for (t in times){
  if (t == "overall"){
    for (ct in celltypes){
      DE_genes_cluster <- filter(pairwise_full, short_cell_type_cluster == ct) %>% select(gene) %>% unique()
      genes_per_cluster[ct, t] <- length(DE_genes_cluster$gene)
    }
  }
  else {
    DE_genes_time <- filter(pairwise_full, time == t)
    genes_per_cluster_time <- as.data.frame(table(DE_genes_time$short_cell_type_cluster))
    for (ct in celltypes){
      if (ct %in% genes_per_cluster_time$Var1){
        genes_per_cluster[ct, t] <- genes_per_cluster_time$Freq[genes_per_cluster_time$Var1==ct]
      }
    }
  }
}
genes_per_cluster[is.na(genes_per_cluster)] <- 0
```

# Identify time at which each cluster has the max DEGs
```{r}
times_list <- c("t1-15HF.t0-UF", "t2-30HF.t0-UF", "t3-60HF.t0-UF", "t4-120HF.t0-UF", "t5-240HF.t0-UF")
genes_per_cluster_max <- genes_per_cluster %>% rowwise() %>% 
  mutate(Top_time = times_list[which.max(c_across(`t1-15HF.t0-UF`:`t5-240HF.t0-UF`))])
genes_per_cluster_max$celltype <- row.names(genes_per_cluster)
```

# Get system info and add to genes per cluster
```{r}
annotation_data <- atlas@meta.data %>% group_by(short_cell_type_cluster) %>% 
  select(c(short_cell_type_cluster, system)) %>% unique()
annotation_data$short_cell_type_cluster <- str_replace_all(annotation_data$short_cell_type_cluster, "_", "-")
genes_per_cluster_max <- left_join(genes_per_cluster_max, annotation_data, by = c("celltype" = "short_cell_type_cluster"))
```

# Get summary table of counts for each system and top time
```{r}
top_time_summary <- genes_per_cluster_max %>% group_by(Top_time, system) %>% summarise(Count = n())
```
```{r}
system_colors <- c("connective tissue" = "#22FAFC", "sensory" = "#FB00FB", "musculoskeletal" = "#E6E90D", "digestive" = "#4B4D53", "transporter" = "#EE2200", "nervous" = "#267CFE", "integumentary" = "#FF7EAD", "cardiovascular" = "#00FB7E", "respiratory" = "#F5960D", "neural crest" = "#DCA3FB", "unknown" = "#FFDCBB", "endocrine" = "#407C0D", "immune" = "#0D85BA", "pronephric" = "#7C420D")
time_colors <- c("t1-15HF.t0-UF" = "orange", "t2-30HF.t0-UF" = "yellow2", "t3-60HF.t0-UF" = "lightgreen", "t4-120HF.t0-UF" = "steelblue1", "t5-240HF.t0-UF" = "mediumpurple1")
```

# Plot with systems on x axis and colors by time
```{r, fig.width=3.5, fig.height=2.25}
most_responsive_time_plot <-  ggplot(top_time_summary, aes(x = system, y = Count, fill = Top_time)) +
  geom_bar(position = "fill", stat = "identity") + theme_minimal() + 
  labs(x = "System", y = "Proportion of Clusters", fill = "Comparison with\nmost DEGs")+
  theme(axis.text.x = element_text(angle = 45, hjust = 1), text = element_text(family = "sans", size = 8), legend.key.size = unit(0.125, "in")) +
  scale_fill_manual(values = time_colors)
most_responsive_time_plot
```
# Save plot
```{r}
save_plot("260624_Pairwise_most_responsive_time_by_system_plot.png", most_responsive_time_plot, base_height = 2.25, base_width = 3)
save_plot("260624_Pairwise_most_responsive_time_by_system_plot.svg", most_responsive_time_plot, base_height = 2.25, base_width = 3)
```

# Session info
```{r}
sessionInfo()
```

