---
title: "Creb3l3 bulk RNA-seq analysis and figures"
author: "Cat Brown"
output:
  html_notebook:
    toc: yes
    toc_float: yes
---

# Load libraries
```{r}
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(tidyverse)
library(writexl)
library(readxl)
library(cowplot)
library(svglite)
library(pheatmap)
library(ggVennDiagram)
library(DESeq2)
```

# Read in Creb3l3 larval bulk RNA-seq data
```{r}
fc_cutoff <- 1.5
p_cutoff <- 0.05

tpmData <- read.csv('C:/Users/cat14/OneDrive - Johns Hopkins/Farber_Lab/Atlas paper/Figure Folders for Server-Lab/Fig 6 - creb3l3/Larval_salmon.merged.gene_tpm.tsv', 
                           header = TRUE, 
                           sep = "\t", 
                           row.names = "gene_id")
countData <- read.csv('C:/Users/cat14/OneDrive - Johns Hopkins/Farber_Lab/Atlas paper/Figure Folders for Server-Lab/Fig 6 - creb3l3/salmon.merged.gene_counts.tsv', 
                           header = TRUE, 
                           sep = "\t", 
                           row.names = "gene_id")
```

# Reshape tpm data
```{r}
tpmData$ensembl_gene_id <- rownames(tpmData)
row.names(tpmData) <- NULL

tpmData <- tpmData %>% 
  pivot_longer(cols = c("mut_hf1", "mut_hf2", "mut_hf3", 
                        "mut_nf1", "mut_nf2", "mut_nf3",
                        "wt_hf1", "wt_hf2", "wt_hf3", 
                        "wt_nf1", "wt_nf2", "wt_nf3",),
               names_to = "sample_id",
               values_to = "TPM") %>%
  mutate(replicate = case_when(grepl("1$", sample_id) ~ 1,
                               grepl("2$", sample_id) ~ 2,
                               grepl("3$", sample_id) ~ 3),
         Diet = case_when(grepl("hf", sample_id) ~ "hf",
                          grepl("nf", sample_id) ~ "nf"),
         genotype = case_when(grepl("mut", sample_id) ~ "mut",
                              grepl("wt", sample_id) ~ "wt"))
```

# Set up metadata for DESeq2
```{r}
cnames <- colnames(countData)
diet <- c("hf", "hf", "hf", "nf","nf","nf", "hf", "hf", "hf", "nf","nf","nf")
genotype <- c("mut","mut","mut","mut","mut","mut","wt","wt","wt","wt","wt","wt") 
replicate <-  c(1,2,3,1,2,3,1,2,3,1,2,3)
columnData <- tibble(cnames, diet, genotype, replicate) %>%
  column_to_rownames(var = "cnames")
```

# DESeq for main effect of genotype
```{r}
dds <- DESeqDataSetFromMatrix(countData = round(countData),
                              colData = columnData,
                              design = ~ diet + genotype)
dds$genotype <- relevel(dds$genotype, ref = "wt")
dds$diet <- relevel(dds$diet, ref = "nf")
dds <- DESeq(dds, test = "LRT", reduced = ~ diet)

res <- results(dds, tidy = TRUE)
colnames(res)[1] <- "ensembl_gene_id"
```
# Get gene names from tpm data and add to results
```{r}
gene_info <- dplyr::select(tpmData, gene_name, ensembl_gene_id) %>% unique()
res <- left_join(res, gene_info, by = "ensembl_gene_id")
```

# Get just significant results
```{r}
sig_res <- filter(res, padj <=0.05)
```

# Save DE results
```{r}
write.csv(res, "260723_creb3l3_bulk_DE_genotype_main_effect.csv")
write.csv(sig_res, "260723_creb3l3_bulk_DE_genotype_main_effect_sig_hits.csv")
```

# Heatmap of count Data for these genes
## Make matrix of counts of relvant genes
```{r}
count_matrix <- countData %>% rownames_to_column(var = "ensembl_gene_id") %>% 
  filter(ensembl_gene_id %in% sig_res$ensembl_gene_id) %>% left_join(gene_info) %>%
  select(!ensembl_gene_id) %>% column_to_rownames(var = "gene_name") %>%
  as.matrix()
```

# Use a max divided normalization instead of the z-score
```{r}
max_exp_df<- as.data.frame(count_matrix) %>% 
  rowwise() %>% summarize(Maximum = max(c_across(1:12)))
creb_matrix_max_divided <- as.data.frame(count_matrix) / unlist(max_exp_df)
creb_matrix_max_divided <- as.matrix(creb_matrix_max_divided)
```

# Plot it
```{r, fig.height=8, fig.width=3.5}
h1 <- pheatmap(creb_matrix_max_divided, cluster_cols = FALSE, annotation_col = columnData, scale = "none", cutree_rows = 3, fontsize = 6, treeheight_row = 4, show_colnames = FALSE)
```
# Get the groupings
```{r}
creb_groups = cutree(h1$tree_row,3)
creb_groups = data.frame(creb_groups)
rownames(creb_groups) = rownames(creb_matrix_max_divided)
```

# Name the groupings
```{r}
creb_groups <- creb_groups %>% mutate(Creb3l3_deletion_response  = case_when(
  creb_groups == 1 ~ "Upregulated",
  creb_groups == 2 ~ "Uniformly Downregulated",
  creb_groups == 3 ~ "Fed Downregulated"
))
```

# Session info
```{r}
sessionInfo()
```

