---
title: "Files for comparison to Zeituni data"
author: "Cat Brown"
output:
  html_notebook:
    toc: yes
    toc_float: yes
---

# Load libraries
```{r}
.libPaths("/apps/R/4.3.2/seurat/5.1.0")
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(tidyverse)
library(writexl)
library(readxl)
library(cowplot)
library(DESeq2)
library(svglite, lib="/home/biederman/R/x86_64-pc-linux-gnu-library/4.3/")
```

# Load atlas
```{r}
atlas <- readRDS("/mnt/sequence/home/cbrown/atlas_paper/260212_WT_atlas_v2_cell_cycle.rds")
```

# Load pairwise DEGs
# Read in DE genes
```{r}
pairwise_full <- read.csv("260129_short_cell_type_cluster_v2_0_pairwise_DE_from_250113.csv")
```

# Filter to comparisions to unfed
```{r}
pairwise_full <- filter(pairwise_full, time %in% c("t1-15HF.t0-UF", "t2-30HF.t0-UF", "t3-60HF.t0-UF", "t4-120HF.t0-UF", "t5-240HF.t0-UF"))
```

# Get just the subset of cells I care about
```{r}
Zeituni_set <- subset(atlas, (tissue %in% c("enteric", "intestine", "digestive general", "liver", "bile duct", "exocrine pancreas", "endocrine pancreas", "esophagus")) | (tree_clusters == "99"))
```

# Get list of short_cell_type_clusters that matter
```{r}
Zeituni_cell_types <- unique(Zeituni_set$short_cell_type_cluster) %>% 
  str_replace_all("_", "-")
```

# Get DE results for just the Zeituni cell types
```{r}
Zeituni_DEGs <- filter(pairwise_full, short_cell_type_cluster %in% Zeituni_cell_types)
```

# Save
```{r}
write_csv(Zeituni_DEGs, "260402_significant_pairwise_to_unfed_DEGs_for_clusters_in_Zeituni.csv")
```

# Pseudo-bulk the entire set
```{r}
pseudo_ct <- AggregateExpression(Zeituni_set, assays = "RNA", return.seurat = T, group.by = c("condition", "sample_name"))
```
# Check naming of cells
```{r}
head(Cells(pseudo_ct))
```
# Run DE across all times (LRT)
```{r}
dds <- DESeqDataSetFromMatrix(countData = pseudo_ct@assays$RNA$counts, 
                              colData = pseudo_ct@meta.data,
                              design = ~condition)
dds$condition <- relevel(dds$condition, ref = "t0-UF")
dds <- DESeq(dds, test = "LRT", useT = TRUE, minmu = 1e-6, minReplicatesForReplace = Inf, reduced = ~1)
res <- results(dds, alpha = 0.05)
res <- rownames_to_column(data.frame(res), "gene")
sig_res <- filter(res, padj <= 0.05)
```
```{r}
write_csv(res, "260402_Zeituni_cell_types_bulk_LRT_results_full.csv")
write_csv(sig_res, "260402_Zeituni_cell_types_bulk_LRT_results_filtered.csv")
```

# Run pairwise DE analysis
```{r}
times = c("t1-15HF.t0-UF", "t2-30HF.t0-UF", "t3-60HF.t0-UF", "t4-120HF.t0-UF", "t5-240HF.t0-UF")
celltype_DE =NULL
for (t in times){
  tryCatch({
    t1=strsplit(t, "[.]")[[1]][1]
    t2=strsplit(t, "[.]")[[1]][2]
    sample = subset(pseudo_ct, (condition %in% c(t1, t2)))
    dds <- DESeqDataSetFromMatrix(countData = sample@assays$RNA$counts, 
                                  colData = sample@meta.data,
                                design = ~condition)
    dds$condition <- relevel(dds$condition, ref = t2)
    dds <- DESeq(dds, test = "LRT", useT = TRUE, minmu = 1e-6, minReplicatesForReplace = Inf, reduced = ~1)
    res <- results(dds, alpha = 0.05)
    res <- rownames_to_column(data.frame(res), "gene")
    res$time <- t
    celltype_DE = rbind(celltype_DE, res)},
    error = function(err){
      bad_ct_time_combos <- c(bad_ct_time_combos, paste(celltype, t, sep = "_"))
    }
  )
}
filtered_celltype_DE <- filter(celltype_DE, padj <= 0.05)
```

```{r}
write_csv(celltype_DE, "260402_Zeituni_cell_types_bulk_pairwise_results_full.csv")
write_csv(filtered_celltype_DE, "260402_Zeituni_cell_types_bulk_pairwise_results_filtered.csv")
```

# Save the seurat object
```{r}
saveRDS(Zeituni_set, "260402_Zeituni_comparison_seurat_file.rds")
```


# Session info
```{r}
sessionInfo()
```