---
title: "MB4-MB26v2.0 propeller stacked bar and cluster tree"
author: "Michelle Biederman"
date: "6/3/25"
output:
  html_notebook:
    toc: yes
    toc_float: yes
---

# Purpose

See if I can combine the cluster tree with the stacked bar plot by putting stacked bar plot in cluster tree order

# Load libpaths

```{r}
.libPaths("/apps/R/4.3.2/seurat/5.1.0")
```

# Load the libraries

```{r, message=FALSE}
library(speckle, lib="/home/biederman/R/x86_64-pc-linux-gnu-library/4.3/")
library(Seurat)
library(CellBench, lib="/home/biederman/R/x86_64-pc-linux-gnu-library/4.3/")
library(limma, lib="/home/biederman/R/x86_64-pc-linux-gnu-library/4.3/")
library(ggplot2)
library(scater, lib="/apps/R/4.3.2/osca/")
library(patchwork)
library(edgeR, lib="/apps/R/4.3.2/osca/")
library(statmod, lib="/apps/R/4.3.2/osca/")
library(dplyr)
library(ggplot2)
library(cowplot)
library(tidyr)
library(ape, lib="/apps/R/4.3.2/DESeq2")
library(scales)
library(stringr)
library(svglite, lib="/home/biederman/R/x86_64-pc-linux-gnu-library/4.3/")
```

# Load data

```{r}
MB.combined <- readRDS("/mnt/sequence/home/biederman/250522_atlas_v2_0.rds")
MB.combined
```

# Check metadata

```{r}
head(MB.combined@meta.data)
```

# Add metadata columns for propeller

```{r}
MB.combined$cluster <- MB.combined$tree_clusters
MB.combined$sample <- MB.combined$sample_name
MB.combined$group <- MB.combined$condition
```

# check new metadata

```{r}
head(MB.combined@meta.data)
```

# Change Ident to cell_type_cluster

```{r}
Idents(MB.combined) <- "cluster"
```

# Perform arcsin square root transformation

```{r}
MB.prop.arcsin <- propeller(MB.combined, transform="asin")
MB.prop.arcsin
```

# Filter out pvalue > 0.05

```{r}
filter(MB.prop.arcsin,P.Value <= 0.05)
```

# Filter out false discovery rates > 0.05

```{r}
filter(MB.prop.arcsin,FDR <= 0.05)
```

Another option for running propeller is for the user to supply the cluster, 
sample and group information explicitly to the propeller function.

# Visualise the results

In the r BiocStyle::Biocpkg("speckle") package there is a plotting function 
plotCellTypeProps that takes a Seurat or SingleCellExperiment object, 
extracts sample and cluster information and outputs a barplot of cell type 
proportions between the samples. The user also has the option of supplying the
cluster and sample cell information instead of an R object. The output is a 
ggplot2 object that the user can then manipulate however they please.

# Get cell type proportions and transformed proportions directly

Alternatively, we can obtain the cell type proportions and transformed 
proportions directly by running the getTransformedProps function which takes 
the cluster and sample information as input. The output from 
getTransformedProps is a list with the cell type counts, transformed 
proportions and proportions as elements.

# reorder factors

```{r}
tree_factors <- rev(c("1", "2", "3", "4", "5", "6", "7", "8", "9", "10", "11", "12", "13", "14", "15", "16", "17", "18", "19", "20", "21", "22", "23", "24", "25", "26", "27", "28", "29", "30", "31", "32", "33", "34", "35", "36", "37", "38", "39", "40", "41", "42", "43", "44", "45", "46", "47", "48", "49", "50", "51", "52", "53", "54", "55", "56", "57", "58", "59", "60", "61", "62", "63", "64", "65", "66", "67", "68", "69", "70", "71", "72", "73", "74", "75", "76", "77", "78", "79", "80", "81", "82", "83", "84", "85", "86", "87", "88", "89", "90", "91", "92", "93", "94", "95", "96", "97", "98", "99", "100", "101", "102", "103", "104", "105", "106", "107", "108", "109", "110", "111", "112", "113", "114", "115", "116", "117", "118", "119", "120", "121", "122", "123", "124", "125", "126", "127", "128", "129", "130", "131", "132", "133", "134", "135", "136", "137", "138", "139", "140", "141", "142", "143", "144", "145", "146", "147", "148", "149", "150", "151", "152", "153", "154", "155", "156", "157", "158", "159", "160", "161", "162", "163", "164", "165", "166", "167", "168", "169", "170", "171", "172", "173", "174", "175", "176", "177", "178", "179", "180", "181", "182", "183", "184", "185", "186", "187", "188", "189", "190", "191", "192", "193", "194", "195", "196", "197", "198", "199", "200", "201", "202", "203", "204", "205", "206", "207", "208", "209", "210", "211", "212", "213", "214", "215", "216", "217", "218", "219", "220", "221", "222", "223", "224", "225", "226", "227", "228", "229", "230", "231", "232", "233", "234", "235", "236", "237", "238", "239", "240", "241", "242", "243", "244", "245", "246", "247", "248", "249", "250", "251", "252", "253", "254", "255", "256", "257", "258", "259", "260", "261", "262", "263", "264", "265", "266", "267", "268", "269", "270", "271"))
```

```{r}
MB.combined$cluster <- as.factor(MB.combined$cluster)
levels(MB.combined$cluster) <- tree_factors
levels(MB.combined$cluster)
```

# colors from ggplot default color palatte

```{r}
colors <- rev(hue_pal()(271))
```

# stacked bar chart for cluster tree 

```{r,fig.height=57,fig.width=10}
par(mar = c(10, 4, 2, 11))
props.asin <- getTransformedProps(MB.combined$cluster, MB.combined$sample, transform="asin")
barplot(props.asin$Proportions, col = colors,legend.text =TRUE, args.legend = list(x = "topleft", inset = c(-0.1,0)),
        ylab="Proportions", las = 2)
#save.image(file="250611_stacked_bar_v2_0.svg")
#dev.copy(svg, "250611_stacked_bar_v2_0.svg", width = 570, height = 100)
#dev.off()
```


# See which models best fit data variances

If you are interested in seeing which models best fit the data in terms of the
cell type variances, there are two plotting functions that can do this: 
plotCellTypeMeanVar and plotCellTypePropsMeanVar. For this particular 
dataset it isn't very informative because there are only three cell "types" 
and no biogical variability.


```{r,fig.height=5, fig.width=10}
props.arcsin <- getTransformedProps(MB.combined$cluster, MB.combined$sample, transform="asin")
par(mfrow=c(1,1))
plotCellTypeMeanVar(props.arcsin$Counts)
plotCellTypePropsMeanVar(props.arcsin$Counts)
```

# Fitting linear models using the transformed proportions directly

If you are like me, you won't feel very comfortable with a black-box approach 
where one function simply spits out a table of results. If you would like to 
have more control over your linear model and include extra covariates then you 
can fit a linear model in a more direct way using the transformed proportions 
that can be obtained by running the getTransformedProps function.

We have already obtained the proportions and transformed proportions when we ran
the getTransformedProps function above. This function outputs a list object 
with three elements: Counts, TransformedProps and Proportions. These are 
all matrices with clusters/cell types in the rows and samples in the columns.

# examine transformed props

```{r}
names(props.arcsin)
head(props.arcsin$TransformedProps)
```

First we need to set up our sample information in much the same way we would if 
we were analysing bulk RNA-seq data. We can pretend that we have pairing 
information which corresponds to our original vs bootstrapped samples to make 
our model a little more complicated for demonstration purposes. 

# set up sample info

```{r}
group <- c("t0_UF","t0_UF","t0_UF","t1_15HF","t1_15HF","t1_15HF","t2_30HF","t2_30HF","t3_60HF","t3_60HF","t3_60HF","t4_120HF","t4_120HF","t5_240HF","t5_240HF")
sampleMB <- c(1,2,3,1,2,3,1,2,1,2,3,1,2,1,2)
mb.df <- data.frame(group,sampleMB)
mb.df
```

# Design model matrix

We can set up a design matrix taking into account group and pairing information.
Please note that the way that propeller has been designed is such that the 
group information is *always* first in the design matrix specification, and 
there is NO intercept. If you are new to design matrices and linear modelling, I
would highly recommend reading the r BiocStyle::Biocpkg("limma") manual, which
is incredibly extensive and covers a variety of different experimental set ups.

```{r}
design <- model.matrix(~ 0 + group + sampleMB, data = mb.df, na.action)
design
```

#ANOVA 

In our example, we have three groups, 10x, CelSeq and DropSeq. In order to 
perform an ANOVA to test for cell type composition differences between these
3 technologies, we can use the propeller.anova function. The coef argument
tells the function which columns of the design matrix correspond to the groups 
we are interested in testing. Here we are interested in the first three columns.

## ANOVA arcsin

```{r}
MB.anova.arcsin <- propeller.anova(prop.list=props.arcsin, design=design, coef = c(1,2,3,4,5,6), 
                robust=TRUE, trend=FALSE, sort=TRUE)
MB.anova.arcsin
```

## ANOVA arcsin with p >0.05 filtered out

```{r}
MB.anova.arcsin %>% filter(P.Value <= 0.05)
```

## ANOVA arcsin with FDR >0.05 filtered out

```{r}
MB.anova.arcsin %>% filter(FDR <= 0.05)
```
Note that the p-values are smaller here than before because we have specified
a pairing vector that states which samples were bootstrapped and which are the 
original samples.

If we were interested in testing only 10x versus DropSeq we could alternatively
use the propeller.ttest function and specify a contrast that tests for this
comparison with our design matrix.

# Plot cluster tree with ape and no node labels

```{r}
MB.combined <- BuildClusterTree(object = MB.combined, reduction = "umap", verbose = TRUE, reorder = TRUE)
```

```{r, fig.height=40, fig.width=8}
# pull the tree
data.tree <- Tool(object = MB.combined, slot = "BuildClusterTree")
data.tree$tip.label <- str_sub(data.tree$tip.label,2,-1)
# plot the tree
tree_plot <- ape::plot.phylo(x = data.tree, direction = "rightwards", type = "tidy")
tree_plot
# had to plot it in the console and export the svg 4000 x 800
```

# Save tree

```{r}
tree_plot
```



# session info

```{r}
sessionInfo()
```




