---
title: "LRT GO and KEGG with lipid timing"
author: "Cat Brown"
output:
  html_notebook:
    toc: yes
    toc_float: yes
---

# Load libraries
```{r, message=FALSE}
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(tidyverse)
library(writexl)
library(readxl)
library(cowplot)
library(svglite)
library(pheatmap)
library(ggpattern)
``` 



# Read in LRT results
```{r}
LRT <- read.csv("Z:/Labs/Farber_Lab/Farber_Lab_Public/snRNA-seq/atlas paper figures/Sup Data - DEGs LRT and pairwise/251211_DESeq_LRT_all_times_filtered_DE_results.csv")
```

# Read in Atlas
```{r}
atlas <- readRDS("C:/Users/cat14/OneDrive - Johns Hopkins/Farber_Lab/Basecaller contents/atlas_paper/260212_WT_atlas_v2_cell_cycle.rds")
```

# Read in lipid timings
```{r}
lipid_timings <- read_xlsx("C:/Users/cat14/OneDrive/Documents/Farber_lab_local_code/260605_lipid_timing_by_cluster.xlsx")
```

# Read in response group info
```{r}
response_groups <- read.csv("C:/Users/cat14/OneDrive - Johns Hopkins/Farber_Lab/Basecaller contents/atlas_paper/260219_LRT_response_groups.csv") 
```

# Get unique LRT genes
```{r}
genes <- LRT %>% select("gene") %>% unique() %>% unlist()
```

# Get average expression levels for all genes in all clusters
```{r}
avg_exp <- AverageExpression(atlas, group.by = c("short_cell_type_cluster", "sample_name"), assay = "RNA", features = genes)
```

# Get system info
```{r}
annotation_data <- atlas@meta.data %>% group_by(short_cell_type_cluster) %>% 
  select(c(short_cell_type_cluster, system)) %>% unique()
annotation_data$short_cell_type_cluster <- str_replace_all(annotation_data$short_cell_type_cluster, "_", "-")
```

# Remove atlas for space
```{r}
rm(atlas)
```

# Restructure avg_exp so that I can combine with LRT
```{r}
avg_exp_df <- as.data.frame(avg_exp[["RNA"]]) %>% 
  rownames_to_column(var = "gene") %>%
  pivot_longer(cols = 2:4023, values_to = "expression", names_to = "cluster_sample") %>%
  separate_wider_delim(cols = "cluster_sample", delim = "_", names = c("cluster", "sample")) %>%
  pivot_wider(names_from = "sample", values_from = "expression")
```

# Combine to filter the average expression to relvant rows
```{r}
LRT_tags <- select(LRT, c("gene", "celltype"))
heatmap_data <- left_join(LRT_tags, avg_exp_df, by = c("gene" = "gene", "celltype" = "cluster"))
```

# Create matrix with gene_cluster as rownames and expression as the data
```{r}
heatmap_matrix <- unite(heatmap_data, "gene", "celltype", col="gene_celltype", sep = "_") %>% unique() %>% relocate(`t3-sample9-60HF`, .after = `t2-sample8-30HF`)  %>% remove_rownames() %>% column_to_rownames(var = "gene_celltype") %>% as.matrix()
```

```{r}
time_colors <- c("unfed" = "hotpink1",  "15min" = "orange", "30min" = "yellow2", "1h" = "lightgreen", "2h" = "steelblue1", "4h" = "mediumpurple1")
sample_times_info <- c("unfed", "unfed", "unfed", "15min", "15min", "15min", "30min", "30min", "1h", "1h", "1h", "2h", "2h", "4h", "4h")
response_colors <- c("1" = "#e41a1c", "2" = "#377eb8", "3" = "#4daf4a", "4" = "#984ea3", "5" = "#ff7f00", "6" = "#ffff33")
sample_times <- data.frame(condition = sample_times_info, row.names = colnames(heatmap_matrix))
```
```{r}
response_groups = column_to_rownames(response_groups, var = "gene_cluster")
```
```{r}
response_groups$response_groups = as.character(response_groups$response_groups)
```

# Figuring out lipid flux details
```{r}
response_groups <- rownames_to_column(response_groups, var = "gene_cluster") %>% 
  separate_wider_delim(cols = "gene_cluster", delim = "_", names = c("gene", "cluster"))
```
```{r}
lipid_timings$cell_type_cluster <- str_replace_all(lipid_timings$cell_type_cluster, "_", "-")
response_lipid <- left_join(response_groups, lipid_timings, by = c("cluster" = "cell_type_cluster"))
```
# Make column on whether it could be flux
```{r}
response_lipid <- response_lipid %>% mutate(Mechanism = case_when(
  Lipid_timing == "t1-15HF" ~ "Flux Possible",
  (Lipid_timing == "t3-60HF") & (response_groups %in% c("3", "5")) ~ "Flux Possible",
  (Lipid_timing == "t5-240HF") & (response_groups == "3") ~ "Flux Possible",
  .default = "Signaling"
))
```
# Reformat for using with heatmap
```{r}
response_lipid <- response_lipid %>% unite(gene, cluster, col = "gene_cluster", sep = "_") %>%
  column_to_rownames(var = "gene_cluster")
```

```{r}
time_colors_2 <- c("t1-15HF" = "orange", "t3-60HF" = "lightgreen", "t5-240HF" = "mediumpurple1")
mech_colors <- c("Flux Possible" = "maroon4", "Signaling" = "lightgray")
```


# Make heatmap
```{r, fig.height=10, fig.width=4}
h1 = pheatmap(heatmap_matrix, scale = "row",  cluster_cols = FALSE, show_rownames = FALSE, cuttree_rows = 6, annotation_col = sample_times, annotation_colors = list(condition = time_colors, response_groups = response_colors, Lipid_timing = time_colors_2, Mechanism = mech_colors), annotation_row = response_lipid, treeheight_row = 8, fontsize = 8)
```

# Summarize systems by response group
# Get the gene and cluster information out of the response group info
```{r}
response_group_data <- rownames_to_column(response_lipid, var = "rowname") %>%
  separate_wider_delim(cols = "rowname", delim = "_", names = c("gene", "cluster"))
```

# Add things for GO analysis
```{r, message=FALSE}
library(clusterProfiler)
library(enrichplot)
library(org.Dr.eg.db)
```

# GO analysis for response groups
```{r}
response_group_GO <- compareCluster(gene~response_groups+Mechanism, data = response_group_data, fun = "enrichGO", OrgDb = org.Dr.eg.db, 
                         keyType = "SYMBOL")
```

# Plot it
```{r, fig.height=9, fig.width=3.5}
response_group_GO_plot <- dotplot(response_group_GO, x = "response_groups", showCategory = 10, label_format = function(x) stringr::str_wrap(x, width=38)) + 
  labs(x = "Response Group") + facet_wrap(~Mechanism) + theme_minimal() +
  scale_size(range = c(1, 3.5)) + set_enrichplot_color(type = 'fill', transform = 'log10') +
  theme(text = element_text(size = 7, family = "sans"), axis.title = element_text(size = 8), 
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
response_group_GO_plot
```
```{r}
save_plot("260608_LRT_response_group_GO_plot.png", response_group_GO_plot, base_height = 9, base_width = 3.5)
save_plot("260608_LRT_response_group_GO_plot.svg", response_group_GO_plot, base_height = 9, base_width = 3.5)
```
# Save GO output
```{r}
saveRDS(response_group_GO, "260608_atlas_paper_LRT_response_group_mechanism_GO_result.rds")
```

# Change the really small p-values so that we can see something in GO
```{r}
response_group_GO@compareClusterResult$p.adjust[response_group_GO@compareClusterResult$p.adjust<(1e-10)] <- 1e-10
```

# Plot it
```{r, fig.height=9, fig.width=3.5}
response_group_GO_plot_2 <- dotplot(response_group_GO, x = "response_groups", showCategory = 10, label_format = function(x) stringr::str_wrap(x, width=38)) + 
  labs(x = "Response Group") + facet_wrap(~Mechanism) + theme_minimal() +
  scale_size(range = c(1, 3.5)) + set_enrichplot_color(type = 'fill', transform = 'log10') +
  theme(text = element_text(size = 7, family = "sans"), axis.title = element_text(size = 8), 
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
response_group_GO_plot_2
```
```{r}
save_plot("260608_LRT_response_group_GO_plot_p_cutoff.png", response_group_GO_plot_2, base_height = 9, base_width = 3.5)
save_plot("260608_LRT_response_group_GO_plot_p_cutoff.svg", response_group_GO_plot_2, base_height = 9, base_width = 3.5)
```
# Convert names
```{r}
ENTREZID <- bitr(response_group_data$gene, fromType = "SYMBOL", toType = "ENTREZID", OrgDb = "org.Dr.eg.db")
ENTREZID <- ENTREZID[order(ENTREZID$SYMBOL), ]
ENTREZID <- ENTREZID[!duplicated(ENTREZID$SYMBOL), ]
response_group_data$ENTREZID <- ENTREZID$ENTREZID[match(response_group_data$gene, ENTREZID$SYMBOL)]
```

```{r}
response_group_KEGG <- compareCluster(ENTREZID~response_groups+Mechanism, data = response_group_data, fun = "enrichKEGG", organism = "dre")
```

```{r, fig.height=9, fig.width=3.5}
response_group_KEGG_plot <- dotplot(response_group_KEGG, x = "response_groups", showCategory = 10, label_format = function(x) stringr::str_wrap(x, width=38)) + 
  labs(x = "Response Group") + facet_wrap(~Mechanism) + theme_minimal() +
  scale_size(range = c(1, 3.5)) + set_enrichplot_color(type = 'fill', transform = 'log10') +
  theme(text = element_text(size = 7, family = "sans"), axis.title = element_text(size = 8), 
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
response_group_KEGG_plot
```
```{r}
save_plot("260608_LRT_response_group_KEGG_plot.png", response_group_KEGG_plot, base_height = 9, base_width = 3.5)
save_plot("260608_LRT_response_group_KEGG_plot.svg", response_group_KEGG_plot, base_height = 9, base_width = 3.5)
```
# Save KEGG output
```{r}
saveRDS(response_group_KEGG, "260608_atlas_paper_LRT_response_group_mechanism_KEGG_result.rds")
```

# Change the really small p-values so that we can see something in GO
```{r}
response_group_KEGG@compareClusterResult$p.adjust[response_group_KEGG@compareClusterResult$p.adjust<(1e-10)] <- 1e-10
```

```{r, fig.height=9, fig.width=3.5}
response_group_KEGG_plot_2 <- dotplot(response_group_KEGG, x = "response_groups", showCategory = 10, label_format = function(x) stringr::str_wrap(x, width=38)) + 
  labs(x = "Response Group") + facet_wrap(~Mechanism) + theme_minimal() +
  scale_size(range = c(1, 3.5)) + set_enrichplot_color(type = 'fill', transform = 'log10') +
  theme(text = element_text(size = 7, family = "sans"), axis.title = element_text(size = 8), 
        axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))
response_group_KEGG_plot_2
```

```{r}
save_plot("260608_LRT_response_group_KEGG_plot_p_cutoff.png", response_group_KEGG_plot_2, base_height = 9, base_width = 3.5)
save_plot("260608_LRT_response_group_KEGG_plot_p_cutoff.svg", response_group_KEGG_plot_2, base_height = 9, base_width = 3.5)
```
# Session info
```{r}
sessionInfo()
```

