---
title: "DE figure updates/fixes"
author: "Cat Brown"
output:
  html_notebook:
    toc: yes
    toc_float: yes
---

# Load libraries
```{r, message=FALSE}
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(tidyverse)
library(writexl)
library(readxl)
library(cowplot)
library(svglite)
library(pheatmap)
library(ggVennDiagram)
library(aplot)
library(RColorBrewer)
``` 
```{r, message=FALSE}
library(ggmagnify)
```


# REad in LRT results
```{r}
LRT <- read.csv("Z:/Labs/Farber_Lab/Farber_Lab_Public/snRNA-seq/atlas paper figures/Sup Data - DEGs LRT and pairwise/251211_DESeq_LRT_all_times_filtered_DE_results.csv")
```

# Get counts of clusters for each gene
```{r}
gene_Cluster_info <- LRT %>% group_by(gene) %>% summarise(Count = n())
```

# plot the histogram
```{r, fig.width=3.25, fig.height=1.75}
gene_count_hist = ggplot(gene_Cluster_info, aes(x = Count)) + geom_histogram(binwidth = 1, boundary = 0) +
  labs(x = "Number of Clusters a LRT DEG is Identified In", y = "Number of Genes") + theme_minimal() + 
  theme(text = element_text(family = "sans", size = 7))
hist_inset <- gene_count_hist + theme(plot.background = element_rect(fill = "white", colour = NA), axis.title = element_blank(), panel.grid.minor = element_blank()) + scale_x_continuous(breaks = c(50, 100, 150)) 
hist_with_inset <- gene_count_hist + geom_magnify(from = c(xmin = 40, xmax = 150, ymin = 0.1, ymax = 5), to = c(xmin = 29, xmax = 152, ymin = 300, ymax = 2300), axes = "y", plot = hist_inset) + scale_y_sqrt(breaks = c(0, 10, 50, 100, 500, 1000, 2000), labels = c(0, 10, 50, 100, 500, 1000, 2000)) 
hist_with_inset
```
```{r}
save_plot("260623_Cluster_genes_histogram_sqrt_scale.png", hist_with_inset, base_height = 1.75, base_width = 3.25)
save_plot("260623_Cluster_genes_histogram_sqrt_scale.svg", hist_with_inset, base_height = 1.75, base_width = 3.25)
```

# Finding good example genes for each response group
## Read in atlas
```{r}
atlas <- readRDS("C:/Users/cat14/OneDrive - Johns Hopkins/Farber_Lab/Basecaller contents/atlas_paper/260212_WT_atlas_v2_cell_cycle.rds")
```

```{r}
response_groups <- read_csv("C:/Users/cat14/OneDrive - Johns Hopkins/Farber_Lab/Basecaller contents/atlas_paper/260219_LRT_response_groups.csv")
```
## I can get genes for all clusters from EnteroAnte-2 so I'm going to subset that out
```{r}
aEC2 <- subset(atlas, short_cell_type_cluster == "EnteroAnte_2")
```

# Making a figure of these 6 options - in ggviolin so I have more control
## Make the data frame
```{r}
vln_df_nupr = data.frame(expression = aEC2[["RNA"]]@data["nupr1b",], condition = aEC2$condition, gene = "nupr1b")
vln_df_flvcr1 = data.frame(expression = aEC2[["RNA"]]@data["flvcr1",], condition = aEC2$condition, gene = "flvcr1")
vln_df_apoa4a = data.frame(expression = aEC2[["RNA"]]@data["apoa4a",], condition = aEC2$condition, gene = "apoa4a")
vln_df_lpin2 = data.frame(expression = aEC2[["RNA"]]@data["lpin2",], condition = aEC2$condition, gene = "lpin2")
vln_df_nocta = data.frame(expression = aEC2[["RNA"]]@data["nocta",], condition = aEC2$condition, gene = "nocta")
vln_df_slc26a6.1 = data.frame(expression = aEC2[["RNA"]]@data["slc26a6.1",], condition = aEC2$condition, gene = "slc26a6.1")
vln_df_examples <- rbind(vln_df_nupr, vln_df_flvcr1, vln_df_apoa4a, vln_df_lpin2, vln_df_nocta, vln_df_slc26a6.1)
```
## Make the gene a factor with the correct levels
```{r}
vln_df_examples$gene <- factor(vln_df_examples$gene, levels = c("nupr1b", "flvcr1", "apoa4a", "lpin2", "nocta", "slc26a6.1"))
```

```{r, fig.width=2.75, fig.height=5.5}
examples_vln <- ggplot(vln_df_examples, aes(x = condition, y = expression)) + 
  geom_violin(aes(fill = condition), trim=TRUE, scale = "width") + 
  geom_jitter(color = "black", size = 0.01) + facet_wrap(~gene, ncol = 1) +
  labs(y = "Expression Level") +
  scale_x_discrete(name = "Condition") + scale_fill_manual(values = c("hotpink1","orange","yellow2", "lightgreen",  "steelblue1", "mediumpurple1")) + theme_minimal() + theme(legend.position = "none", text = element_text(size = 7, family = "sans"), axis.text.x = element_text(angle=45, hjust = 1), strip.text = element_text(hjust = 0.5, face = "bold", size = 8), axis.title = element_text(size = 8))
examples_vln
```
```{r}
save_plot("260623_LRT_response_group_example_violins.png", examples_vln, base_height = 5.5, base_width = 2.75)
save_plot("260623_LRT_response_group_example_violins.svg", examples_vln, base_height = 5.5, base_width = 2.75)
```

# Session info
```{r}
sessionInfo()
```

