---
title: "MB4-26_denovo_RPCA_25mt_M3MADupc_M3MADgpc_200gpc_v1_corrected"
author: "Michelle Biederman"
date: "6/9/23"
output:
  html_notebook:
    toc: yes
    toc_float: yes
---

# Purpose

6-15-23: All products were based on 173 PCs.  UMAP had previously been created with dimenstions 1:153.  All other workflow steps (ScaleData, RunPCA, FindNeighbors, and FindClusters) used dimensions 1:173.  I made an updated dimplot with UMI set to 173.  All other products remain the same.

6-9-23: on 6-2-23 I didn't properly add nFeature_RNA_M3MAD column to metadata and incorrectly subsetted nFeature_RNA < nCount_RNA_M3MAD instead of nFeature_RNA < nFeature_RNA_M3MAD.  Updated to more recent jackstraw indicating 173 sig PCs.

1. RPCA integrate all samples MB4-26.
  * elbow plot and Jackstraw to determine # of PCs determined from previous work = 153 PCs
2. Cluster at multiple resolutions
  * tissue/organ organization 
  * cell-type organization (benchmarked by germline pre-cursors clustering separately)

QC subset: 
  * < 25% mt
  * < median + 3 MAD UMIs per cell
  * < median + 3 MAD genes per cell
  * > 200 genes per cell
  
Normalization: log norm 10000

Variable feature selection: 2000

[Satija lab vignette - Fast integration using reciprocal PCA (RPCA)](https://satijalab.org/seurat/articles/integration_rpca.html)

[Satija lab vignette - aggregated feature expression by identity class](https://satijalab.org/seurat/reference/aggregateexpression)

# Data preparation

## Loaded .libPaths

```{r}
.libPaths( "/apps/R/4.2.2/seurat/4.3.0")
.libPaths()
```

## Loaded libraries

```{r, message=FALSE}
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(RColorBrewer)
library(cowplot)
library(tidyverse)
library(writexl)
```

## Loaded MB4, MB5, MB6, MB9, MB10, MB11, MB12, MB13, MB14, MB15, MB16, MB23, MB24, MB25, and MB26 Ensembl data

```{r}
query_data_MB4<-Read10X("/mnt/sequence/10x/MB4_count/outs/filtered_feature_bc_matrix")
```

```{r}
query_data_MB5<-Read10X("/mnt/sequence/10x/MB5_count/outs/filtered_feature_bc_matrix")
```

```{r}
query_data_MB6<-Read10X("/mnt/sequence/10x/MB6_count/outs/filtered_feature_bc_matrix")
```

```{r}
query_data_MB9<-Read10X("/mnt/sequence/10x/MB9_count/outs/filtered_feature_bc_matrix")
```

```{r}
query_data_MB10<-Read10X("/mnt/sequence/10x/MB10_count/outs/filtered_feature_bc_matrix")
```

```{r}
query_data_MB11<-Read10X("/mnt/sequence/10x/MB11_count/outs/filtered_feature_bc_matrix")
```

```{r}
query_data_MB12<-Read10X("/mnt/sequence/10x/MB12_count/outs/filtered_feature_bc_matrix")
```

```{r}
query_data_MB13<-Read10X("/mnt/sequence/10x/MB13_count/outs/filtered_feature_bc_matrix")
```

```{r}
query_data_MB14<-Read10X("/mnt/sequence/10x/MB14_count/outs/filtered_feature_bc_matrix")
```

```{r}
query_data_MB15<-Read10X("/mnt/sequence/10x/MB15_count/outs/filtered_feature_bc_matrix")
```

```{r}
query_data_MB16<-Read10X("/mnt/sequence/10x/MB16_count/outs/filtered_feature_bc_matrix")
```

```{r}
query_data_MB23<-Read10X("/mnt/sequence/10x/MB23_count/outs/filtered_feature_bc_matrix")
```

```{r}
query_data_MB24<-Read10X("/mnt/sequence/10x/MB24_count/outs/filtered_feature_bc_matrix")
```

```{r}
query_data_MB25<-Read10X("/mnt/sequence/10x/MB25_count/outs/filtered_feature_bc_matrix")
```

```{r}
query_data_MB26<-Read10X("/mnt/sequence/10x/MB26_count/outs/filtered_feature_bc_matrix")
```

## Created Seurat objects

```{r, message=FALSE, results=FALSE}
query_seurat_MB4 <- CreateSeuratObject( query_data_MB4, project = "MB4" )
query_seurat_MB5 <- CreateSeuratObject( query_data_MB5, project = "MB5" )
query_seurat_MB6 <- CreateSeuratObject( query_data_MB6, project = "MB6" )
query_seurat_MB9 <- CreateSeuratObject( query_data_MB9, project = "MB9" )
query_seurat_MB10 <- CreateSeuratObject( query_data_MB10, project = "MB10" )
query_seurat_MB11 <- CreateSeuratObject( query_data_MB11, project = "MB11" )
query_seurat_MB12 <- CreateSeuratObject( query_data_MB12, project = "MB12" )
query_seurat_MB13 <- CreateSeuratObject( query_data_MB13, project = "MB13" )
query_seurat_MB14 <- CreateSeuratObject( query_data_MB14, project = "MB14" )
query_seurat_MB15 <- CreateSeuratObject( query_data_MB15, project = "MB15" )
query_seurat_MB16 <- CreateSeuratObject( query_data_MB16, project = "MB16" )
query_seurat_MB23 <- CreateSeuratObject( query_data_MB23, project = "MB23" )
query_seurat_MB24 <- CreateSeuratObject( query_data_MB24, project = "MB24" )
query_seurat_MB25 <- CreateSeuratObject( query_data_MB25, project = "MB25" )
query_seurat_MB26 <- CreateSeuratObject( query_data_MB26, project = "MB26" )
```

## Cleaned up

```{r}
rm(query_data_MB4)
rm(query_data_MB6)
rm(query_data_MB5)
rm(query_data_MB9)
rm(query_data_MB10)
rm(query_data_MB11)
rm(query_data_MB12)
rm(query_data_MB13)
rm(query_data_MB14)
rm(query_data_MB15)
rm(query_data_MB16)
rm(query_data_MB23)
rm(query_data_MB24)
rm(query_data_MB25)
rm(query_data_MB26)
gc()
```

## Check objects

```{r}
query_seurat_MB4
query_seurat_MB5
query_seurat_MB6
query_seurat_MB9
query_seurat_MB10
query_seurat_MB11
query_seurat_MB12
query_seurat_MB13
query_seurat_MB14
query_seurat_MB15
query_seurat_MB16
query_seurat_MB23
query_seurat_MB24
query_seurat_MB25
query_seurat_MB26
```

## Added column "sample_name"

In case I want to display results by sample name.  This is not a good order for plotting.
```{r}
query_seurat_MB4@meta.data$sample_name <-query_seurat_MB4@meta.data[1:length(query_seurat_MB4@meta.data),"sample_name"] <- "t0_sample3_UF"
query_seurat_MB5@meta.data$sample_name <-query_seurat_MB5@meta.data[1:length(query_seurat_MB5@meta.data),"sample_name"] <- "t2_sample7_30HF"
query_seurat_MB6@meta.data$sample_name <-query_seurat_MB6@meta.data[1:length(query_seurat_MB6@meta.data),"sample_name"] <- "t0_sample1_UF"
query_seurat_MB9@meta.data$sample_name <-query_seurat_MB9@meta.data[1:length(query_seurat_MB9@meta.data),"sample_name"] <- "t2_sample8_30HF"
query_seurat_MB10@meta.data$sample_name <-query_seurat_MB10@meta.data[1:length(query_seurat_MB10@meta.data),"sample_name"] <- "t3_sample9_60HF"
query_seurat_MB11@meta.data$sample_name <-query_seurat_MB11@meta.data[1:length(query_seurat_MB11@meta.data),"sample_name"] <- "t0_sample2_UF"
query_seurat_MB12@meta.data$sample_name <-query_seurat_MB12@meta.data[1:length(query_seurat_MB12@meta.data),"sample_name"] <- "t4_sample12_120HF"
query_seurat_MB13@meta.data$sample_name <-query_seurat_MB13@meta.data[1:length(query_seurat_MB13@meta.data),"sample_name"] <- "t3_sample10_60HF"
query_seurat_MB14@meta.data$sample_name <-query_seurat_MB14@meta.data[1:length(query_seurat_MB14@meta.data),"sample_name"] <- "t5_sample14_240HF"
query_seurat_MB15@meta.data$sample_name <-query_seurat_MB15@meta.data[1:length(query_seurat_MB15@meta.data),"sample_name"] <- "t4_sample13_120HF"
query_seurat_MB16@meta.data$sample_name <-query_seurat_MB16@meta.data[1:length(query_seurat_MB16@meta.data),"sample_name"] <- "t5_sample15_240HF"
query_seurat_MB23@meta.data$sample_name <-query_seurat_MB23@meta.data[1:length(query_seurat_MB23@meta.data),"sample_name"] <- "t3_sample11_60HF"
query_seurat_MB24@meta.data$sample_name <-query_seurat_MB24@meta.data[1:length(query_seurat_MB24@meta.data),"sample_name"] <- "t1_sample4_15HF"
query_seurat_MB25@meta.data$sample_name <-query_seurat_MB25@meta.data[1:length(query_seurat_MB25@meta.data),"sample_name"] <- "t1_sample5_15HF"
query_seurat_MB26@meta.data$sample_name <-query_seurat_MB26@meta.data[1:length(query_seurat_MB26@meta.data),"sample_name"] <- "t1_sample6_15HF"
```

## Added column "condition"

Because I want my conditions to plot in the right order, I am adding a number to the beginning of each.  This will add complications because I have to use the ` to enclose them because R doesn't like certain terms that start with a number, but I'm doing it anyway, for now.

```{r}
query_seurat_MB4@meta.data$condition <-query_seurat_MB4@meta.data[1:length(query_seurat_MB4@meta.data),"condition"] <- "t0_UF"
query_seurat_MB5@meta.data$condition <-query_seurat_MB5@meta.data[1:length(query_seurat_MB5@meta.data),"condition"] <- "t2_30HF"
query_seurat_MB6@meta.data$condition <-query_seurat_MB6@meta.data[1:length(query_seurat_MB6@meta.data),"condition"] <- "t0_UF"
query_seurat_MB9@meta.data$condition <-query_seurat_MB9@meta.data[1:length(query_seurat_MB9@meta.data),"condition"] <- "t2_30HF"
query_seurat_MB10@meta.data$condition <-query_seurat_MB10@meta.data[1:length(query_seurat_MB10@meta.data),"condition"] <- "t3_60HF"
query_seurat_MB11@meta.data$condition <-query_seurat_MB11@meta.data[1:length(query_seurat_MB11@meta.data),"condition"] <- "t0_UF"
query_seurat_MB12@meta.data$condition <-query_seurat_MB12@meta.data[1:length(query_seurat_MB12@meta.data),"condition"] <- "t4_120HF"
query_seurat_MB13@meta.data$condition <-query_seurat_MB13@meta.data[1:length(query_seurat_MB13@meta.data),"condition"] <- "t3_60HF"
query_seurat_MB14@meta.data$condition <-query_seurat_MB14@meta.data[1:length(query_seurat_MB14@meta.data),"condition"] <- "t5_240HF"
query_seurat_MB15@meta.data$condition <-query_seurat_MB15@meta.data[1:length(query_seurat_MB15@meta.data),"condition"] <- "t4_120HF"
query_seurat_MB16@meta.data$condition <-query_seurat_MB16@meta.data[1:length(query_seurat_MB16@meta.data),"condition"] <- "t5_240HF"
query_seurat_MB23@meta.data$condition <-query_seurat_MB23@meta.data[1:length(query_seurat_MB23@meta.data),"condition"] <- "t3_60HF"
query_seurat_MB24@meta.data$condition <-query_seurat_MB24@meta.data[1:length(query_seurat_MB24@meta.data),"condition"] <- "t1_15HF"
query_seurat_MB25@meta.data$condition <-query_seurat_MB25@meta.data[1:length(query_seurat_MB25@meta.data),"condition"] <- "t1_15HF"
query_seurat_MB26@meta.data$condition <-query_seurat_MB26@meta.data[1:length(query_seurat_MB26@meta.data),"condition"] <- "t1_15HF"
```

# Calculate percent.mt column

```{r}
query_seurat_MB4 <- PercentageFeatureSet(query_seurat_MB4, "^mt-", col.name = "percent.mt")
query_seurat_MB5 <- PercentageFeatureSet(query_seurat_MB5, "^mt-", col.name = "percent.mt")
query_seurat_MB6 <- PercentageFeatureSet(query_seurat_MB6, "^mt-", col.name = "percent.mt")
query_seurat_MB9 <- PercentageFeatureSet(query_seurat_MB9, "^mt-", col.name = "percent.mt")
query_seurat_MB10 <- PercentageFeatureSet(query_seurat_MB10, "^mt-", col.name = "percent.mt")
query_seurat_MB11 <- PercentageFeatureSet(query_seurat_MB11, "^mt-", col.name = "percent.mt")
query_seurat_MB12 <- PercentageFeatureSet(query_seurat_MB12, "^mt-", col.name = "percent.mt")
query_seurat_MB13 <- PercentageFeatureSet(query_seurat_MB13, "^mt-", col.name = "percent.mt")
query_seurat_MB14 <- PercentageFeatureSet(query_seurat_MB14, "^mt-", col.name = "percent.mt")
query_seurat_MB15 <- PercentageFeatureSet(query_seurat_MB15, "^mt-", col.name = "percent.mt")
query_seurat_MB16 <- PercentageFeatureSet(query_seurat_MB16, "^mt-", col.name = "percent.mt")
query_seurat_MB23 <- PercentageFeatureSet(query_seurat_MB23, "^mt-", col.name = "percent.mt")
query_seurat_MB24 <- PercentageFeatureSet(query_seurat_MB24, "^mt-", col.name = "percent.mt")
query_seurat_MB25 <- PercentageFeatureSet(query_seurat_MB25, "^mt-", col.name = "percent.mt")
query_seurat_MB26 <- PercentageFeatureSet(query_seurat_MB26, "^mt-", col.name = "percent.mt")
```

# Calculate Count and Feature MADs median + 3MAD UMIs and *genes per cell* upper thresholds{.tabset}

## MB6

```{r}
query_seurat_MB6@meta.data <- query_seurat_MB6@meta.data %>%
  mutate(nFeature_RNA_M3MAD = median(query_seurat_MB6$nFeature_RNA) + 3*mad(query_seurat_MB6$nFeature_RNA, center = median(query_seurat_MB6$nFeature_RNA), constant = 1.4826, na.rm = FALSE,
    low = FALSE, high = FALSE)) %>%
  mutate(nCount_RNA_M3MAD = median(query_seurat_MB6$nCount_RNA) +3*mad(query_seurat_MB6$nCount_RNA, center = median(query_seurat_MB6$nCount_RNA), constant = 1.4826, na.rm = FALSE,
    low = FALSE, high = FALSE))
```

```{r}
head(query_seurat_MB6@meta.data)
```

## MB11

```{r}
query_seurat_MB11@meta.data <- query_seurat_MB11@meta.data %>%
  mutate(nFeature_RNA_M3MAD = median(query_seurat_MB11$nFeature_RNA) + 3*mad(query_seurat_MB11$nFeature_RNA, center = median(query_seurat_MB11$nFeature_RNA), constant = 1.4826, na.rm = FALSE,
    low = FALSE, high = FALSE)) %>%
  mutate(nCount_RNA_M3MAD = median(query_seurat_MB11$nCount_RNA) +3*mad(query_seurat_MB11$nCount_RNA, center = median(query_seurat_MB11$nCount_RNA), constant = 1.4826, na.rm = FALSE,
    low = FALSE, high = FALSE))
```

```{r}
head(query_seurat_MB11@meta.data)
```

## MB4

```{r}
query_seurat_MB4@meta.data <- query_seurat_MB4@meta.data %>%
  mutate(nFeature_RNA_M3MAD = median(query_seurat_MB4$nFeature_RNA) + 3*mad(query_seurat_MB4$nFeature_RNA, center = median(query_seurat_MB4$nFeature_RNA), constant = 1.4826, na.rm = FALSE,
    low = FALSE, high = FALSE)) %>%
  mutate(nCount_RNA_M3MAD = median(query_seurat_MB4$nCount_RNA) +3*mad(query_seurat_MB4$nCount_RNA, center = median(query_seurat_MB4$nCount_RNA), constant = 1.4826, na.rm = FALSE,
    low = FALSE, high = FALSE))
```

```{r}
head(query_seurat_MB4@meta.data)
```

## MB24

```{r}
query_seurat_MB24@meta.data <- query_seurat_MB24@meta.data %>%
  mutate(nFeature_RNA_M3MAD = median(query_seurat_MB24$nFeature_RNA) + 3*mad(query_seurat_MB24$nFeature_RNA, center = median(query_seurat_MB24$nFeature_RNA), constant = 1.4826, na.rm = FALSE,
    low = FALSE, high = FALSE)) %>%
  mutate(nCount_RNA_M3MAD = median(query_seurat_MB24$nCount_RNA) +3*mad(query_seurat_MB24$nCount_RNA, center = median(query_seurat_MB24$nCount_RNA), constant = 1.4826, na.rm = FALSE,
    low = FALSE, high = FALSE))
```

```{r}
head(query_seurat_MB24@meta.data)
```

## MB25

```{r}
query_seurat_MB25@meta.data <- query_seurat_MB25@meta.data %>%
  mutate(nFeature_RNA_M3MAD = median(query_seurat_MB25$nFeature_RNA) + 3*mad(query_seurat_MB25$nFeature_RNA, center = median(query_seurat_MB25$nFeature_RNA), constant = 1.4826, na.rm = FALSE,
    low = FALSE, high = FALSE)) %>%
  mutate(nCount_RNA_M3MAD = median(query_seurat_MB25$nCount_RNA) +3*mad(query_seurat_MB25$nCount_RNA, center = median(query_seurat_MB25$nCount_RNA), constant = 1.4826, na.rm = FALSE,
    low = FALSE, high = FALSE))
```

```{r}
head(query_seurat_MB25@meta.data)
```

## MB26

```{r}
query_seurat_MB26@meta.data <- query_seurat_MB26@meta.data %>%
  mutate(nFeature_RNA_M3MAD = median(query_seurat_MB26$nFeature_RNA) + 3*mad(query_seurat_MB26$nFeature_RNA, center = median(query_seurat_MB26$nFeature_RNA), constant = 1.4826, na.rm = FALSE,
    low = FALSE, high = FALSE)) %>%
  mutate(nCount_RNA_M3MAD = median(query_seurat_MB26$nCount_RNA) +3*mad(query_seurat_MB26$nCount_RNA, center = median(query_seurat_MB26$nCount_RNA), constant = 1.4826, na.rm = FALSE,
    low = FALSE, high = FALSE))
```

```{r}
head(query_seurat_MB26@meta.data)
```

## MB5

```{r}
query_seurat_MB5@meta.data <- query_seurat_MB5@meta.data %>%
  mutate(nFeature_RNA_M3MAD = median(query_seurat_MB5$nFeature_RNA) + 3*mad(query_seurat_MB5$nFeature_RNA, center = median(query_seurat_MB5$nFeature_RNA), constant = 1.4826, na.rm = FALSE,
    low = FALSE, high = FALSE)) %>%
  mutate(nCount_RNA_M3MAD = median(query_seurat_MB5$nCount_RNA) +3*mad(query_seurat_MB5$nCount_RNA, center = median(query_seurat_MB5$nCount_RNA), constant = 1.4826, na.rm = FALSE,
    low = FALSE, high = FALSE))
```

```{r}
head(query_seurat_MB5@meta.data)
```

## MB9

```{r}
query_seurat_MB9@meta.data <- query_seurat_MB9@meta.data %>%
  mutate(nFeature_RNA_M3MAD = median(query_seurat_MB9$nFeature_RNA) + 3*mad(query_seurat_MB9$nFeature_RNA, center = median(query_seurat_MB9$nFeature_RNA), constant = 1.4826, na.rm = FALSE,
    low = FALSE, high = FALSE)) %>%
  mutate(nCount_RNA_M3MAD = median(query_seurat_MB9$nCount_RNA) +3*mad(query_seurat_MB9$nCount_RNA, center = median(query_seurat_MB9$nCount_RNA), constant = 1.4826, na.rm = FALSE,
    low = FALSE, high = FALSE))
```

```{r}
head(query_seurat_MB9@meta.data)
```

## MB10

```{r}
query_seurat_MB10@meta.data <- query_seurat_MB10@meta.data %>%
  mutate(nFeature_RNA_M3MAD = median(query_seurat_MB10$nFeature_RNA) + 3*mad(query_seurat_MB10$nFeature_RNA, center = median(query_seurat_MB10$nFeature_RNA), constant = 1.4826, na.rm = FALSE,
    low = FALSE, high = FALSE)) %>%
  mutate(nCount_RNA_M3MAD = median(query_seurat_MB10$nCount_RNA) +3*mad(query_seurat_MB10$nCount_RNA, center = median(query_seurat_MB10$nCount_RNA), constant = 1.4826, na.rm = FALSE,
    low = FALSE, high = FALSE))
```

```{r}
head(query_seurat_MB10@meta.data)
```

## MB13

```{r}
query_seurat_MB13@meta.data <- query_seurat_MB13@meta.data %>%
  mutate(nFeature_RNA_M3MAD = median(query_seurat_MB13$nFeature_RNA) + 3*mad(query_seurat_MB13$nFeature_RNA, center = median(query_seurat_MB13$nFeature_RNA), constant = 1.4826, na.rm = FALSE,
    low = FALSE, high = FALSE)) %>%
  mutate(nCount_RNA_M3MAD = median(query_seurat_MB13$nCount_RNA) +3*mad(query_seurat_MB13$nCount_RNA, center = median(query_seurat_MB13$nCount_RNA), constant = 1.4826, na.rm = FALSE,
    low = FALSE, high = FALSE))
```

```{r}
head(query_seurat_MB13@meta.data)
```

## MB23

```{r}
query_seurat_MB23@meta.data <- query_seurat_MB23@meta.data %>%
  mutate(nFeature_RNA_M3MAD = median(query_seurat_MB23$nFeature_RNA) + 3*mad(query_seurat_MB23$nFeature_RNA, center = median(query_seurat_MB23$nFeature_RNA), constant = 1.4826, na.rm = FALSE,
    low = FALSE, high = FALSE)) %>%
  mutate(nCount_RNA_M3MAD = median(query_seurat_MB23$nCount_RNA) +3*mad(query_seurat_MB23$nCount_RNA, center = median(query_seurat_MB23$nCount_RNA), constant = 1.4826, na.rm = FALSE,
    low = FALSE, high = FALSE))
```

```{r}
head(query_seurat_MB23@meta.data)
```

## MB12

```{r}
query_seurat_MB12@meta.data <- query_seurat_MB12@meta.data %>%
  mutate(nFeature_RNA_M3MAD = median(query_seurat_MB12$nFeature_RNA) + 3*mad(query_seurat_MB12$nFeature_RNA, center = median(query_seurat_MB12$nFeature_RNA), constant = 1.4826, na.rm = FALSE,
    low = FALSE, high = FALSE)) %>%
  mutate(nCount_RNA_M3MAD = median(query_seurat_MB12$nCount_RNA) +3*mad(query_seurat_MB12$nCount_RNA, center = median(query_seurat_MB12$nCount_RNA), constant = 1.4826, na.rm = FALSE,
    low = FALSE, high = FALSE))
```

```{r}
head(query_seurat_MB12@meta.data)
```

## MB15

```{r}
query_seurat_MB15@meta.data <- query_seurat_MB15@meta.data %>%
  mutate(nFeature_RNA_M3MAD = median(query_seurat_MB15$nFeature_RNA) + 3*mad(query_seurat_MB15$nFeature_RNA, center = median(query_seurat_MB15$nFeature_RNA), constant = 1.4826, na.rm = FALSE,
    low = FALSE, high = FALSE)) %>%
  mutate(nCount_RNA_M3MAD = median(query_seurat_MB15$nCount_RNA) +3*mad(query_seurat_MB15$nCount_RNA, center = median(query_seurat_MB15$nCount_RNA), constant = 1.4826, na.rm = FALSE,
    low = FALSE, high = FALSE))
```

```{r}
head(query_seurat_MB15@meta.data)
```

## MB14

```{r}
query_seurat_MB14@meta.data <- query_seurat_MB14@meta.data %>%
  mutate(nFeature_RNA_M3MAD = median(query_seurat_MB14$nFeature_RNA) + 3*mad(query_seurat_MB14$nFeature_RNA, center = median(query_seurat_MB14$nFeature_RNA), constant = 1.4826, na.rm = FALSE,
    low = FALSE, high = FALSE)) %>%
  mutate(nCount_RNA_M3MAD = median(query_seurat_MB14$nCount_RNA) +3*mad(query_seurat_MB14$nCount_RNA, center = median(query_seurat_MB14$nCount_RNA), constant = 1.4826, na.rm = FALSE,
    low = FALSE, high = FALSE))
```

```{r}
head(query_seurat_MB14@meta.data)
```

## MB16

```{r}
query_seurat_MB16@meta.data <- query_seurat_MB16@meta.data %>%
  mutate(nFeature_RNA_M3MAD = median(query_seurat_MB16$nFeature_RNA) + 3*mad(query_seurat_MB16$nFeature_RNA, center = median(query_seurat_MB16$nFeature_RNA), constant = 1.4826, na.rm = FALSE,
    low = FALSE, high = FALSE)) %>%
  mutate(nCount_RNA_M3MAD = median(query_seurat_MB16$nCount_RNA) +3*mad(query_seurat_MB16$nCount_RNA, center = median(query_seurat_MB16$nCount_RNA), constant = 1.4826, na.rm = FALSE,
    low = FALSE, high = FALSE))
```

```{r}
head(query_seurat_MB16@meta.data)
```

# Save unfiltered objects

```{r}
saveRDS(query_seurat_MB6,"/mnt/sequence/home/biederman/230602_v1/unfiltered/230602_sample1_UF_MB6_corrected.rds")
saveRDS(query_seurat_MB11,"/mnt/sequence/home/biederman/230602_v1/unfiltered/230602_sample2_UF_MB11_corrected.rds")
saveRDS(query_seurat_MB4,"/mnt/sequence/home/biederman/230602_v1/unfiltered/230602_sample3_UF_MB4_corrected.rds")
saveRDS(query_seurat_MB24,"/mnt/sequence/home/biederman/230602_v1/unfiltered/230602_sample4_15_MB24_corrected.rds")
saveRDS(query_seurat_MB25,"/mnt/sequence/home/biederman/230602_v1/unfiltered/230602_sample5_15_MB25_corrected.rds")
saveRDS(query_seurat_MB26,"/mnt/sequence/home/biederman/230602_v1/unfiltered/230602_sample6_15_MB26_corrected.rds")
saveRDS(query_seurat_MB5,"/mnt/sequence/home/biederman/230602_v1/unfiltered/230602_sample7_30_MB5_corrected.rds")
saveRDS(query_seurat_MB9,"/mnt/sequence/home/biederman/230602_v1/unfiltered/230602_sample8_30_MB9_corrected.rds")
saveRDS(query_seurat_MB10,"/mnt/sequence/home/biederman/230602_v1/unfiltered/230602_sample9_60_MB10_corrected.rds")
saveRDS(query_seurat_MB13,"/mnt/sequence/home/biederman/230602_v1/unfiltered/230602_sample10_60_MB13_corrected.rds")
saveRDS(query_seurat_MB23,"/mnt/sequence/home/biederman/230602_v1/unfiltered/230602_sample11_60_MB23_corrected.rds")
saveRDS(query_seurat_MB12,"/mnt/sequence/home/biederman/230602_v1/unfiltered/230602_sample12_120_MB12_corrected.rds")
saveRDS(query_seurat_MB15,"/mnt/sequence/home/biederman/230602_v1/unfiltered/230602_sample13_120_MB15_corrected.rds")
saveRDS(query_seurat_MB14,"/mnt/sequence/home/biederman/230602_v1/unfiltered/230602_sample14_240_MB14_corrected.rds")
saveRDS(query_seurat_MB16,"/mnt/sequence/home/biederman/230602_v1/unfiltered/230602_sample15_240_MB16_corrected.rds")
```

# Subset < 25% mt, < median + 3 MAD UMIs per cell, < median + 3 MAD genes per cell, > 200 genes per cell

```{r}
query_seurat_MB4_F <- subset(query_seurat_MB4, percent.mt < 25 & nCount_RNA < nCount_RNA_M3MAD & nFeature_RNA < nFeature_RNA_M3MAD & nFeature_RNA > 200)
query_seurat_MB5_F <- subset(query_seurat_MB5, percent.mt < 25 & nCount_RNA < nCount_RNA_M3MAD & nFeature_RNA < nFeature_RNA_M3MAD & nFeature_RNA > 200)
query_seurat_MB6_F <- subset(query_seurat_MB6,percent.mt < 25 & nCount_RNA < nCount_RNA_M3MAD & nFeature_RNA < nFeature_RNA_M3MAD & nFeature_RNA > 200)
query_seurat_MB9_F <- subset(query_seurat_MB9, percent.mt < 25 & nCount_RNA < nCount_RNA_M3MAD & nFeature_RNA < nFeature_RNA_M3MAD & nFeature_RNA > 200)
query_seurat_MB10_F <- subset(query_seurat_MB10,percent.mt < 25 & nCount_RNA < nCount_RNA_M3MAD & nFeature_RNA < nFeature_RNA_M3MAD & nFeature_RNA > 200)
query_seurat_MB11_F <- subset(query_seurat_MB11,percent.mt < 25 & nCount_RNA < nCount_RNA_M3MAD & nFeature_RNA < nFeature_RNA_M3MAD & nFeature_RNA > 200)
query_seurat_MB12_F <- subset(query_seurat_MB12, percent.mt < 25 & nCount_RNA < nCount_RNA_M3MAD & nFeature_RNA < nFeature_RNA_M3MAD & nFeature_RNA > 200)
query_seurat_MB13_F <- subset(query_seurat_MB13, percent.mt < 25 & nCount_RNA < nCount_RNA_M3MAD & nFeature_RNA < nFeature_RNA_M3MAD & nFeature_RNA > 200)
query_seurat_MB14_F <- subset(query_seurat_MB14,percent.mt < 25 & nCount_RNA < nCount_RNA_M3MAD & nFeature_RNA < nFeature_RNA_M3MAD & nFeature_RNA > 200)
query_seurat_MB15_F <- subset(query_seurat_MB15, percent.mt < 25 & nCount_RNA < nCount_RNA_M3MAD & nFeature_RNA < nFeature_RNA_M3MAD & nFeature_RNA > 200)
query_seurat_MB16_F <- subset(query_seurat_MB16,percent.mt < 25 & nCount_RNA < nCount_RNA_M3MAD & nFeature_RNA < nFeature_RNA_M3MAD & nFeature_RNA > 200)
query_seurat_MB23_F <- subset(query_seurat_MB23,percent.mt < 25 & nCount_RNA < nCount_RNA_M3MAD & nFeature_RNA < nFeature_RNA_M3MAD & nFeature_RNA > 200)
query_seurat_MB24_F <- subset(query_seurat_MB24, percent.mt < 25 & nCount_RNA < nCount_RNA_M3MAD & nFeature_RNA < nFeature_RNA_M3MAD & nFeature_RNA > 200)
query_seurat_MB25_F <- subset(query_seurat_MB25, percent.mt < 25 & nCount_RNA < nCount_RNA_M3MAD & nFeature_RNA < nFeature_RNA_M3MAD & nFeature_RNA > 200)
query_seurat_MB26_F <- subset(query_seurat_MB26,percent.mt < 25 & nCount_RNA < nCount_RNA_M3MAD & nFeature_RNA < nFeature_RNA_M3MAD & nFeature_RNA > 200)
```

# Confirm subsets and total filtered nuclei

```{r}
query_seurat_MB4_F
query_seurat_MB5_F
query_seurat_MB6_F
query_seurat_MB9_F
query_seurat_MB10_F
query_seurat_MB11_F
query_seurat_MB12_F
query_seurat_MB13_F
query_seurat_MB14_F
query_seurat_MB15_F
query_seurat_MB16_F
query_seurat_MB23_F
query_seurat_MB24_F
query_seurat_MB25_F
query_seurat_MB26_F
```

# Save filtered objects

```{r}
saveRDS(query_seurat_MB6_F,"/mnt/sequence/home/biederman/230602_v1/filtered/230602_sample1_UF_MB6_25mt_200gpc_M3MADgpc_M3MADupc_corrected.rds")
saveRDS(query_seurat_MB11_F,"/mnt/sequence/home/biederman/230602_v1/filtered/230602_sample2_UF_MB11_25mt_200gpc_M3MADgpc_M3MADupc_corrected.rds")
saveRDS(query_seurat_MB4_F,"/mnt/sequence/home/biederman/230602_v1/filtered/230602_sample3_UF_MB4_25mt_200gpc_M3MADgpc_M3MADupc_corrected.rds")
saveRDS(query_seurat_MB24_F,"/mnt/sequence/home/biederman/230602_v1/filtered/230602_sample4_15_MB24_25mt_200gpc_M3MADgpc_M3MADupc_corrected.rds")
saveRDS(query_seurat_MB25_F,"/mnt/sequence/home/biederman/230602_v1/filtered/230602_sample5_15_MB25_25mt_200gpc_M3MADgpc_M3MADupc_corrected.rds")
saveRDS(query_seurat_MB26_F,"/mnt/sequence/home/biederman/230602_v1/filtered/230602_sample6_15_MB26_25mt_200gpc_M3MADgpc_M3MADupc_corrected.rds")
saveRDS(query_seurat_MB5_F,"/mnt/sequence/home/biederman/230602_v1/filtered/230602_sample7_30_MB5_25mt_200gpc_M3MADgpc_M3MADupc_corrected.rds")
saveRDS(query_seurat_MB9_F,"/mnt/sequence/home/biederman/230602_v1/filtered/230602_sample8_30_MB9_25mt_200gpc_M3MADgpc_M3MADupc_corrected.rds")
saveRDS(query_seurat_MB10_F,"/mnt/sequence/home/biederman/230602_v1/filtered/230602_sample9_60_MB10_25mt_200gpc_M3MADgpc_M3MADupc_corrected.rds")
saveRDS(query_seurat_MB13_F,"/mnt/sequence/home/biederman/230602_v1/filtered/230602_sample10_60_MB13_25mt_200gpc_M3MADgpc_M3MADupc_corrected.rds")
saveRDS(query_seurat_MB23_F,"/mnt/sequence/home/biederman/230602_v1/filtered/230602_sample11_60_MB23_25mt_200gpc_M3MADgpc_M3MADupc_corrected.rds")
saveRDS(query_seurat_MB12_F,"/mnt/sequence/home/biederman/230602_v1/filtered/230602_sample12_120_MB12_25mt_200gpc_M3MADgpc_M3MADupc_corrected.rds")
saveRDS(query_seurat_MB15_F,"/mnt/sequence/home/biederman/230602_v1/filtered/230602_sample13_120_MB15_25mt_200gpc_M3MADgpc_M3MADupc_corrected.rds")
saveRDS(query_seurat_MB14_F,"/mnt/sequence/home/biederman/230602_v1/filtered/230602_sample14_240_MB14_25mt_200gpc_M3MADgpc_M3MADupc_corrected.rds")
saveRDS(query_seurat_MB16_F,"/mnt/sequence/home/biederman/230602_v1/filtered/230602_sample15_240_MB16_25mt_200gpc_M3MADgpc_M3MADupc_corrected.rds")
```

## Cleaned up

```{r}
rm(query_seurat_MB4)
rm(query_seurat_MB6)
rm(query_seurat_MB5)
rm(query_seurat_MB9)
rm(query_seurat_MB10)
rm(query_seurat_MB11)
rm(query_seurat_MB12)
rm(query_seurat_MB13)
rm(query_seurat_MB14)
rm(query_seurat_MB15)
rm(query_seurat_MB16)
rm(query_seurat_MB23)
rm(query_seurat_MB24)
rm(query_seurat_MB25)
rm(query_seurat_MB26)
gc()
```

# combine samples into list

```{r}
MB.list <- c(query_seurat_MB6_F, query_seurat_MB11_F, query_seurat_MB4_F,query_seurat_MB24_F, query_seurat_MB25_F, query_seurat_MB26_F,query_seurat_MB5_F, query_seurat_MB9_F, query_seurat_MB10_F,query_seurat_MB13_F, query_seurat_MB23_F, query_seurat_MB12_F,query_seurat_MB15_F, query_seurat_MB14_F, query_seurat_MB16_F)
```

## Confirm list

```{r}
MB.list
```

# Normalization and Finding Variable Features

After removing unwanted cells from the dataset, the next step is to normalize the data. By default, we employ a global-scaling normalization method “LogNormalize” that normalizes the feature expression measurements for each cell by the total expression, multiplies this by a scale factor (10,000 by default), and log-transforms the result. Normalized values are stored in pbmc[["RNA"]]@data.

We next calculate a subset of features that exhibit high cell-to-cell variation in the dataset (i.e, they are highly expressed in some cells, and lowly expressed in others). We and others have found that focusing on these genes in downstream analysis helps to highlight biological signal in single-cell datasets.

Our procedure in Seurat is described in detail here, and improves on previous versions by directly modeling the mean-variance relationship inherent in single-cell data, and is implemented in the FindVariableFeatures() function. By default, we return 2,000 features per dataset. These will be used in downstream analysis, like PCA.

```{r, message = FALSE, results = FALSE}
for (i in 1:length(MB.list)) {
    MB.list[[i]] <- NormalizeData(MB.list[[i]], verbose = FALSE)
    MB.list[[i]] <- FindVariableFeatures(MB.list[[i]], selection.method = "vst", 
        nfeatures = 2000, verbose = FALSE)
}
```

# select integration features, scale data, Run PCA

features that are repeatedly variable across datasets for integration run PCA on each

```{r}
features <- SelectIntegrationFeatures(object.list = MB.list)
MB.list <- lapply(X = MB.list, FUN = function(x) {
    x <- ScaleData(x, features = features, verbose = FALSE)
    x <- RunPCA(x, features = features, verbose = FALSE)
})
```

# RPCA integration

## Find integration Anchors

```{r}
MB.anchors <- FindIntegrationAnchors(object.list = MB.list, anchor.features = features, reduction = "rpca")
```

## Integrate

```{r}
MB.combined <- IntegrateData(anchorset = MB.anchors)
```

## Clean up

```{r}
rm(MB.anchors)
rm(MB.list)
gc()
```

# Change to "integrated" assay

specify that we will perform downstream analysis on the corrected data note that the
original unmodified data still resides in the 'RNA' assay

```{r}
DefaultAssay(MB.combined) <- "integrated"
```

# Check integrated object

```{r}
MB.combined
```

# Run the standard workflow for visualization and clustering

regressed out %mt and UMIs
```{r}
MB.combined <- ScaleData(MB.combined, vars.to.regress = c("percent.mt","nCount_RNA"), verbose = FALSE)
MB.combined <- RunPCA(MB.combined, npcs = 173, verbose = FALSE)
MB.combined <- RunUMAP(MB.combined, reduction = "pca", dims = 1:173)
MB.combined <- FindNeighbors(MB.combined, reduction = "pca", dims = 1:173)
```

# Double check # PCs to include

300 PCs seems like a good place to start.

## Elbow plot

```{r}
ElbowPlot(MB.combined, ndims = 300, reduction = "pca")
```

### Double check Jackstraw plot (optional)

Made a jackstraw plot like in the Farnsworth paper with my 300 PCs
```{r, message=FALSE, results=FALSE}
#MB.combined <- JackStraw(MB.combined, reduction = "pca", dims = 200, num.replicate = 25, verbose = TRUE)
```

```{r}
#MB.combined <- ScoreJackStraw(MB.combined, dims = 1:200)
```

Plotted jackstraw
```{r, fig.height = 20, fig.width = 10}
#JackStrawPlot(MB.combined, reduction = "pca", dims = 1:200) + theme(legend.position = "bottom")
```

### Save integrated .rds

```{r}
saveRDS(MB.combined, file = "/mnt/sequence/home/biederman/230602_v1/filtered/230602_25mt_200gpc_M3MADgpc_M3MADupc_RPCA_integrated_corrected.rds")
```

# Try UMAPs of different resolutions{.tabset}

## Resolution = 20

```{r, message = FALSE, results=FALSE}
DefaultAssay(MB.combined) <- "integrated"
MB.combined <- FindClusters(MB.combined, resolution = 20, verbose = TRUE)
MB.combined <- RunUMAP(MB.combined, reduction = "pca", dims = 1:173)
```

```{r, fig.width=10, fig.height=16}
p1 <- DimPlot(MB.combined, reduction = "umap", group.by = "seurat_clusters", label=TRUE, label.size = 4, repel = TRUE, raster=FALSE) + theme(legend.position = "bottom")
p1
```

# save plot

```{r, fig.width=10, fig.height=16}
save_plot("230602_v1_RPCA_integrated_res_20_UMAP.png", p1, base_height = 10, base_width = 5, )
```

# table

```{r}
Cells <- table(MB.combined$seurat_clusters)
xnames <- names(Cells)
Percent <- round(((prop.table(Cells))*100),2)
MB.table <- cbind(Cells,Percent)
MB.table
```

# save rds

```{r}
saveRDS(MB.combined, file = "/mnt/sequence/home/biederman/230602_v1/filtered/230602_25mt_200gpc_M3MADgpc_M3MADupc_RPCA_integrated_corrected.rds")
```

# Aggregate expression

```{r}
agg_res_20 <- AggregateExpression(object = MB.combined, assays = "RNA", group.by = "seurat_clusters")
genes<-rownames(agg_res_20$RNA)
agg_res_20 <- as.data.frame(agg_res_20, row.names = genes)
agg_res_20 <- cbind(genes,agg_res_20)
head(agg_res_20)
```

# write aggregated expression Excel

```{r}
write_xlsx(agg_res_20, path = "/mnt/sequence/home/biederman/230602_v1/filtered/230602_25mt_200gpc_M3MADgpc_M3MADupc_RPCA_integrated_agg_res_20_corrected.xlsx")
```

# find markers for every cluster compared to all remaining cells, report only the positive ones

```{r}
DefaultAssay(MB.combined) <- "RNA"
MB.markers <- FindAllMarkers(MB.combined, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)
MB.markers %>%
    group_by(cluster) %>%
    top_n(n = 20, wt = avg_log2FC) -> top20
#top20
```

# write top20 Excel

```{r}
write_xlsx(top20, path = "/mnt/sequence/home/biederman/230602_v1/filtered/230602_25mt_200gpc_M3MADgpc_M3MADupc_RPCA_integrated_res_20_top20_corrected.xlsx")
write_xlsx(MB.markers, path = "/mnt/sequence/home/biederman/230602_v1/filtered/230602_25mt_200gpc_M3MADgpc_M3MADupc_RPCA_integrated_res_20_markers_corrected.xlsx")
```

# write gene set csv for cellxgene

```{r}
top20_csv <- as_tibble(top20) %>% 
  select(cluster,gene)
top20_csv$gene_set_name <- top20_csv$cluster
top20_csv$gene_symbol <- top20_csv$gene
top20_csv <- add_column(top20_csv, gene_set_description = NA,
             gene_description = NA)
top20_csv <- top20_csv %>% select(gene_set_name, gene_set_description, gene_symbol, gene_description)
top20_csv
write.csv(top20_csv, file = "/mnt/sequence/home/biederman/230602_v1/filtered/230602_25mt_200gpc_M3MADgpc_M3MADupc_RPCA_integrated_res_20_top20_marker_gene-set_corrected.csv", row.names = FALSE, fileEncoding = "ASCII")
```

# Session info

```{r}
sessionInfo()
```