---
title: "Annotation v2.0"
author: "Cat Brown and Michelle Biederman"
date: "5/22/25"
output:
  html_notebook:
    toc: yes
    toc_float: yes
---

# Purpose

10-27-25: Cluster tree added from 250603_stacked_bar_cluster_tree.Rmd

5-22-25: Now all relative to cluster tree order!  Clusters are cluster tree order, but I kept the old seurat metadata names, too.

5-19-25: v1.5 update based on 250113_v1.4_annotation_update_short_cell_types_and_DE_for_all_cell_types.Rmd

Code from Cat Brown home/cbrown/241118_atlas_v1_3_creation_and_DE.Rmd

8-16-24 changed to updated atlas v1.1 annotations 9-23-24 Cat updated 240923_atlas_v1_2_creation_and_DE.Rmd http://basecaller.ciwemb.edu/home/cbrown/240923_atlas_v1_2_creation_and_DE.Rmd 11-18-24 Cat updated 241118_atlas_v1_3_creation_and_DE.Rmd http://basecaller.ciwemb.edu/home/cbrown/241118_atlas_v1_3_creation_and_DE.Rmd 1-13-25 Updating to v1.4 with short cell type annotations (10 or fewer characters, each) Adding metadata part is based on Michelle’s code in 240814_v1_1_all_annotations_and_metadata.Rmd. DEseq code I previously wrote and run on v1.2

10-16-25: Added ape library and rebuild of cluster tree

# Load libPaths

```{r}
.libPaths("/apps/R/4.3.2/seurat/5.1.0")
```

# Load libraries

```{r, message = false}
library(dplyr)
library(Seurat)
library(patchwork)
library(ggplot2)
library(tidyverse)
library(writexl)
library(readxl)
library(cowplot)
library(DESeq2)
library(svglite, lib="/home/biederman/R/x86_64-pc-linux-gnu-library/4.3/")
library(ape, lib="/apps/R/4.3.2/DESeq2")
```

# Load initial atlas

```{r}
MB.combined <- readRDS("/mnt/sequence/home/biederman/230602_v1/filtered/230602_25mt_200gpc_M3MADgpc_M3MADupc_RPCA_integrated_corrected.rds")
MB.combined
head(MB.combined@meta.data)
```

# Load annotations

```{r}
cellAnnotations<-read.csv("/mnt/sequence/home/biederman/250522_Annotations_v2.csv", header = TRUE, sep=",")
cellAnnotations
``` 

# Add annotations

```{r}
MB.combined@meta.data$tree_clusters<-cellAnnotations$Cluster[as.numeric(MB.combined@meta.data$seurat_clusters)]
MB.combined@meta.data$system<-cellAnnotations$system[as.numeric(MB.combined@meta.data$seurat_clusters)]
MB.combined@meta.data$tissue<-cellAnnotations$tissue[as.numeric(MB.combined@meta.data$seurat_clusters)]
MB.combined@meta.data$cell_type_general<-cellAnnotations$cell_type_general[as.numeric(MB.combined@meta.data$seurat_clusters)]
MB.combined@meta.data$cell_type_cluster<-cellAnnotations$cell_type_cluster[as.numeric(MB.combined@meta.data$seurat_clusters)]
MB.combined@meta.data$short_cell_type<-cellAnnotations$short_cell_type[as.numeric(MB.combined@meta.data$seurat_clusters)]
MB.combined@meta.data$short_cell_type_cluster<-cellAnnotations$short_cell_type_cluster[as.numeric(MB.combined@meta.data$seurat_clusters)]
MB.combined@meta.data$seurat_cell_type_cluster<-cellAnnotations$seurat_cell_type_cluster[as.numeric(MB.combined@meta.data$seurat_clusters)]
MB.combined@meta.data$seurat_short_cell_type_cluster<-cellAnnotations$seurat_short_cell_type_cluster[as.numeric(MB.combined@meta.data$seurat_clusters)]
MB.combined@meta.data
```

```{r}
MB.combined@meta.data$tree_clusters <- as.factor(MB.combined@meta.data$tree_clusters)
MB.combined@meta.data
```

# Rebuild tree with v2.0

```{r}
DefaultAssay(MB.combined) <- "integrated"
Idents(MB.combined) <- "tree_clusters"
MB.combined <- BuildClusterTree(object = MB.combined, reduction = "umap", verbose = TRUE, reorder = TRUE)
```

# Plot cluster tree with ape and no node labels

```{r, fig.height=40, fig.width=8}
# pull the tree
data.tree <- Tool(object = MB.combined, slot = "BuildClusterTree")
data.tree$tip.label <- str_sub(data.tree$tip.label,2,-1)
# plot the tree
tree_plot <- ape::plot.phylo(x = data.tree, direction = "rightwards", type = "tidy")
tree_plot
# had to plot it in the console and export the svg 4000 x 800
```

# Save updated atlas

```{r}
saveRDS(MB.combined, "250522_atlas_v2_0.rds")
```

# New tree_clusters UMAP

```{r,fig.height=20, fig.width=20}
plot <- DimPlot(MB.combined, reduction = "umap", group.by = "tree_clusters", label=TRUE, label.size = 4, repel = TRUE, raster=TRUE) + theme(legend.position = "none")
plot
```

# save plot

```{r}
save_plot("250522_atlas_tree_clusters_v2_0.png", plot, base_height = 20, base_width = 20)
save_plot("250522_atlas_short_tree_clusters_v2_0.svg", plot, base_height = 20, base_width = 20)
```

# Cluster tree UMAP rastered with legend

```{r,fig.height=20, fig.width=30}
plot <- DimPlot(MB.combined, reduction = "umap", group.by = "tree_clusters", label=TRUE, label.size = 4, repel = TRUE, raster=TRUE) + theme(legend.position = "right")
plot
```

# save plot

```{r}
save_plot("250522_atlas_tree_clusters_legend_v2_0.png", plot, base_height = 20, base_width = 20)
save_plot("250522_atlas_short_tree_clusters_legend_v2_0.svg", plot, base_height = 20, base_width = 20)
```

# New short cell type cluster UMAP

```{r,fig.height=20, fig.width=20}
plot <- DimPlot(MB.combined, reduction = "umap", group.by = "short_cell_type_cluster", label=TRUE, label.size = 3.3, repel = TRUE, raster=TRUE) + theme(legend.position = "none")
plot
```

# save plot

```{r}
save_plot("250522_atlas_short_celltype_cluster_v2_0.png", plot, base_height = 20, base_width = 20)
save_plot("250522_atlas_short_celltype_cluster_v2_0.svg", plot, base_height = 20, base_width = 20)
```

```{r}
Idents(MB.combined) <- "tree_clusters"
levels(MB.combined)
```


# New short cell type cluster UMAP

```{r,fig.height=40, fig.width=20}

plot <- DimPlot(MB.combined, reduction = "umap", group.by = "short_cell_type_cluster", label=TRUE, label.size = 3.2, repel = TRUE, raster=TRUE) + theme(legend.position = "bottom")
plot
```

# New short cell type general UMAP

```{r,fig.height=20, fig.width=20}
plot <- DimPlot(MB.combined, reduction = "umap", group.by = "short_cell_type", label=TRUE, label.size = 4, repel = TRUE, raster=TRUE) + theme(legend.position = "none")
plot
```

# save plot

```{r}
save_plot("250522_atlas_short_celltype_general_v2_0.png", plot, base_height = 20, base_width = 20)
save_plot("250522_atlas_short_celltype_general_v2_0.svg", plot, base_height = 20, base_width = 20)
```

# New short cell type cluster UMAP

```{r,fig.height=20, fig.width=20}
plot <- DimPlot(MB.combined, reduction = "umap", group.by = "short_cell_type_cluster", label=TRUE, label.size = 3.4, repel = TRUE, raster=TRUE) + theme(legend.position = "none")
plot
```

# New tissue UMAP with legend

```{r, fig.height=20, fig.width=18}
plot <- DimPlot(MB.combined, reduction = "umap", group.by = "tissue", raster=TRUE) +
  theme(legend.position = "bottom") +
  theme(legend.text=element_text(size=18)) +
  labs(title = "Major Systems") +
  theme(plot.title = element_text(size=22))
plot
```

# New tissue UMAP

```{r, fig.height=20, fig.width=20}
plot2 <- DimPlot(MB.combined, reduction = "umap", group.by = "tissue", label=TRUE, label.size = 6, repel = TRUE, raster=TRUE) + theme(legend.position = "none") +
  labs(title = "Tissues") +
  theme(plot.title = element_text(size=22))
plot2
```

# save plot

```{r}
save_plot("250522_atlas_tissue_v2_0_legend.png", plot, base_height = 20, base_width = 20)
save_plot("250522_atlas_tissue_v2_0_nolegend.png", plot2, base_height = 20, base_width = 20)
```

# New system UMAP with legend

```{r, fig.height=20, fig.width=18}
plot <- DimPlot(MB.combined, reduction = "umap", group.by = "system", raster=TRUE) +
  theme(legend.position = "bottom") +
  theme(legend.text=element_text(size=18)) +
  labs(title = "Major Systems") +
  theme(plot.title = element_text(size=22))
plot
```

# New system UMAP without legend 

```{r, fig.height=20, fig.width=20}
plot2 <- DimPlot(MB.combined, reduction = "umap", group.by = "system", label=TRUE, label.size = 6, repel = TRUE, raster=TRUE) +
  theme(legend.position = "none") +
  theme(legend.text=element_text(size=18)) +
  labs(title = "Major Systems") +
  theme(plot.title = element_text(size=22))
plot2
```

# save plot

```{r}
save_plot("250522_atlas_system_v2_0_legend.png", plot, base_height = 20, base_width = 20)
save_plot("250522_atlas_system_v2_0_nolegend.png", plot2, base_height = 20, base_width = 20)
```

# New short cell type UMAP

```{r, fig.height=20, fig.width=20}
plot <- DimPlot(MB.combined, reduction = "umap", group.by = "short_cell_type", label=TRUE, label.size = 5, repel = TRUE, raster=FALSE) + theme(legend.position = "none") +
  labs(title = "Cell Types", size = 24) +
  theme(plot.title = element_text(size=22))
plot
```

# save plot

```{r}
save_plot("250522_atlas_short_celltype__v2_0.png", plot, base_height = 20, base_width = 20)
```

# Session Info

```{r}
sessionInfo()
```